### 2014-11-18 poster/poster.tex
model spec table

%==============================================================================
%== template for LATEX poster =================================================
%==============================================================================
%
%--A0 beamer slide-------------------------------------------------------------
\documentclass[final]{beamer}
\usepackage[orientation=portrait,size=a0,
            scale=1.25         % font scale factor
           ]{beamerposter}

\geometry{
  hmargin=2.5cm, % little modification of margins
}

%
\usepackage[utf8]{inputenc}

\linespread{1.15}
%
%==The poster style============================================================
\usetheme{sharelatex}

%==Title, date and authors of the poster=======================================
\title
[CSSS 501 MLE Poster Session, 17 November 2014 -- http://github.com/pschmied/bikecounts-mle] % Conference
{ % Poster title
Weather and Seasonal Factors Affecting Daily Bicycle Counts at Fremont St. Bridge, Seattle
}

\author{ % Authors
Peter Schmiedeskamp\inst{1}, Weiran Zhao\inst{1}
}
\institute
[University of Washington] % General University
{
\inst{1} University of Washington, Seattle
}
\date{\today}

\begin{document}
\begin{frame}[t]
%==============================================================================
\begin{multicols}{3}
%==============================================================================
%==The poster content==========================================================
%==============================================================================

\section{Introduction \& Research Questions}
This project develops a statistical model for predicting the volume of
bicycle traffic per day on the Fremont bridge, in order to better
understand factors believed to influence likelihood of bicycling.
These predictors include weather factors such as precipitation and
temperature, season, day of the week, as well as consideration of a
general trend over time.

\section{Data}
In late 2012, the Seattle Department of Transportation deployed an
automated bicycle detection system that has been collecting bicycle
counts since its installation. These data were downloaded from the
City of Seattle's public data portal, aggregated by day, and then
joined by date to historic weather data queried from Forecast.io. In
this analysis, the collection period starts on October 31, 2012 and
ends October 30, 2014, for a total of 730 daily observations.

These data are summarized graphically in figures \ref{fg:hist} and
\ref{fg:timeseries} below. From the histogram, we see what looks to be
a Poisson distribution with over dispersion. From the timeseries plot,
we see that there is a clear yearly pattern, as well as a weekly
pattern.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/dp1}
\caption{Histogram of bicycle counts}
\label{fg:hist}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/dp2}
\caption{Counts by day}
\label{fg:timeseries}
\end{center}
\end{figure}

\section{Model Specification}
Because some over dispersion was evident in the counts, we chose to
fit a negative binomial model. This model included the following variables

\begin{table}[htdp]
\begin{center}
\begin{tabular}{ll}
  \hline\hline
  Variable & Description \\
  \hline
  Count* & Number of bicycles / day \\
  Daylight hours & Length of daylight in hours \\
  Holiday & Was the day a holiday? \\
  UW & Was University of Washington in session? \\
  Max temp & Maximum temperature for the day \\
  Max precip & Daily max precip (inches) in any hour \\
  Day number & Sequentially numbered day of study \\
  Sat--Fri & Dummy variables for day of week \\
  \hline\hline
    & \multicolumn{1}{r}{*dependent variable}
\end{tabular}
\end{center}
\label{default}
\end{table}%

\section{Model Fit}

\section{Simulated Counterfactual Results}

\subsection{Temperature}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c1}
\caption{Simulated effect of daily maximum temperature on bicycle counts}
\label{fg:temp}
\end{center}
\end{figure}

\subsection{Precipitation}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c2}
\caption{Simulated effect of daily precipitation (in inches) on bicycle counts}
\label{fg:maxtemp}
\end{center}
\end{figure}

\subsection{Day of the Week}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c3}
\caption{Simulated bicycle counts differences by day of week}
\label{fg:dow}
\end{center}
\end{figure}

\subsection{Seasonality}
In considering how to address seasonality, we were concerned that
Seattle's Pacific Maritime Climate does not reflect a traditional four
season calendar year. We felt that hours of sunshine and whether or
not it was the ``school season'' (i.e. the University of Washington
was in session) were better approximations of season for Seattle.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c4}
\caption{Effect of season (measured continuously as number of daylight hours)}
\label{default}
\end{center}
\end{figure}

\section{Conclusions}
%==============================================================================
%==End of content==============================================================
%==============================================================================

%--References------------------------------------------------------------------

\subsection{References}

\begin{thebibliography}{99}

\bibitem{ref1} J.~Doe, Article name, \textit{Phys. Rev. Lett.}

\bibitem{ref2} J.~Doe, J. Smith, Other article name, \textit{Phys. Rev. Lett.}

\bibitem{web} \url{http://www.google.pl}

\end{thebibliography}
%--End of references-----------------------------------------------------------

\end{multicols}

%==============================================================================
\end{frame}
\end{document}

### 2014-11-18 R/nbinom.R
AVP plot

@@ -10,6 +10,7 @@ library(MASS)
 library(pscl)
 
 
+
 ## Load the saved dataset
 wb <- read.csv("data/weatherbike2yr.csv",
                header=T, stringsAsFactors=FALSE)
@@ -62,7 +63,7 @@ getyhat <- function(model) {
   # Takes a model object as created by glm, generates yhats
   x <- model.matrix(model)[,-1]
   pe <- model$coefficients
-  1/(1+exp(-cbind(rep(1,nrow(x)),x)%*%pe))
+  exp(cbind(rep(1,nrow(x)),x)%*%pe) 
 }
 
 binys <- function(y, yhat, by=0.05) {
@@ -107,6 +108,13 @@ mod1_sb <- simbetas(mod1_m)             # Simulate our betas
 
 ## Model fit ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
+## Actual versus predicted plot
+avpp1 <- ggplot(data.frame(actual=wb$count, predicted=getyhat(mod1_m)),
+                aes(x=actual, y=predicted)) +
+         geom_abline(intercept=0, slope=1, color="red") +
+         geom_point() + theme_bw()
+
+ggsave(file="avpp1.pdf", path="fig", width=7, height=7, units=("in"))
 
 
 ## Counterfactual simulations of model 1 ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

### 2014-11-18 R/nbinom.R
add general trend, shrink fig heights

@@ -130,7 +130,7 @@ m1c1 <- ggplot(mod1_tidy1, aes(x=temperatureMax, y=pe, ymin=lower,
         geom_line() + geom_ribbon(alpha=0.3) +
         xlab("Daily Max Temp") + ylab("Bicycles") + theme_bw()
 
-ggsave(file="m1c1.pdf", path="fig", width=7, height=5, units=("in"))
+ggsave(file="m1c1.pdf", path="fig", width=7, height=4, units=("in"))
 
 
 ## cf2 - Effect of max precipitation
@@ -144,7 +144,7 @@ m1c2 <- ggplot(mod1_tidy2, aes(x=precipIntensityMax, y=pe, ymin=lower,
         geom_line() + geom_ribbon(alpha=0.3) +
         xlab("Daily Max Precip (inches)") + ylab("Bicycles") + theme_bw()
 
-ggsave(file="m1c2.pdf", path="fig", width=7, height=5, units=("in"))
+ggsave(file="m1c2.pdf", path="fig", width=7, height=4, units=("in"))
 
 
 ## cf3 - Effect of day of the week
@@ -171,7 +171,7 @@ m1c3 <- ggplot(mod1_tidy3, aes(x=Day, y=pe, ymin=lower,
         geom_point() + geom_errorbar(width=0.2) +
         xlab("Day of the week") + ylab("Bicycles") + theme_bw()
 
-ggsave(file="m1c3.pdf", path="fig", width=7, height=5, units=("in"))
+ggsave(file="m1c3.pdf", path="fig", width=7, height=4, units=("in"))
 
 
 ## cf4 - Effect of season (measured continuously via daylight hours)
@@ -185,4 +185,17 @@ m1c4 <- ggplot(mod1_tidy4, aes(x=daylighth, y=pe, ymin=lower,
         geom_line() + geom_ribbon(alpha=0.3) +
         xlab("Daylight (hours)") + ylab("Bicycles") + theme_bw()
 
-ggsave(file="m1c4.pdf", path="fig", width=7, height=5, units=("in"))
+ggsave(file="m1c4.pdf", path="fig", width=7, height=4, units=("in"))
+
+
+## cf5 - General trend over time
+mod1_cf5 <- cfMake2(mod1_f, wb, X=1:730)
+mod1_res5 <- loglinsimev(mod1_cf5, mod1_sb) # Run the sim
+mod1_tidy5 <- fortify.longsim(mod1_res5, mod1_cf5)
+
[…]

### 2014-11-18 poster/poster.tex
draft 1 of poster

@@
 { % Poster title
-Weather and Seasonal Factors Affecting Daily Bicycle Counts at Fremont St. Bridge, Seattle
+  Weather and Seasonal Factors Affecting Daily Bicycle Counts at
+  Fremont St. Bridge, Seattle %
 }
@@
 \section{Model Specification}
-Because some over dispersion was evident in the counts, we chose to
-fit a negative binomial model. This model included the following variables
+Because over dispersion was evident in the counts, we chose to fit a
+negative binomial model. This model includes the following variables,
+and was fit using the \texttt{pscl::glm.nb()} function in R.
 
@@
   Daylight hours & Length of daylight in hours \\
-  Holiday & Was the day a holiday? \\
+  Holiday & Was the day a holiday (recognized by UW)? \\
   UW & Was University of Washington in session? \\
@@
 
-\section{Model Fit}
+\section{Model Fit and Diagnostics}
+
+As the actual-versus-predicted plot below shows, the model is a
+reasonably good fit. However, we note several outliers falling far
+from the $y=x$ line, perhaps indicating some other unaccounted source
+of variation in the dataset.
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.15\textwidth]{../fig/avpp1}
+\caption{Actual versus predicted values from model}
+\label{fg:avp}
+\end{center}
+\end{figure}
 
@@
 \subsection{Temperature}
+
+As expected, the effect of temperature seems clear. Bicyclists at this
+location appear to prefer warmer weather universally over cooler
+weather. In a less temperate climate, we might expect bicycle counts
+to fall again in very high temperature conditions.
 
@@
 \subsection{Precipitation}
+Similiarly, we see the strong negative effect of precipitation in our simulated
+results.
 
@@
 \includegraphics[width=0.3\textwidth]{../fig/m1c2}
-\caption{Simulated effect of daily precipitation (in inches) on bicycle counts}
+\caption{Simulated effect of daily precipitation (in inches) on
+  bicycle counts}
 \label{fg:maxtemp}
@@
 \subsection{Day of the Week}
-
+As identified in our descriptive timeseries plot, these simulated
+results confirm the presense of a weekly pattern in bicycling, with
+much higher numbers of bicyclists during the week, falling off
+somewhat toward Friday, and dropping precipitously on the weekends.
 \begin{figure}[htbp]
@@
 \includegraphics[width=0.3\textwidth]{../fig/m1c4}
-\caption{Effect of season (measured continuously as number of daylight hours)}
-\label{default}
+\caption{Effect of season (measured continuously as number of daylight
+  hours)}
+\label{fg:daylight}
+\end{center}
+\end{figure}
+
+\subsection{General trend over time}
+In our model we included a term that was a sequential numbering of the
+days of the study period as to model a general trend over time. In the
+visualization of the simulated results, we do see a general increase
+in bicycle counts over this two year span.
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.3\textwidth]{../fig/m1c5}
+\caption{Simulated bicycle counts over study period}
+\label{fg:trend}
 \end{center}
@@
 \section{Conclusions}
+Over all, this model appears to be reasonably
+predictive, with the results confirming our intuitions about factors
+influencing bicycling rates. One major limitation, is that we still do
+not account for the possibility that there is some timeseries
+autoregression ocurring on a day-to-day basis. This is a problematic
+limitation, in that we believe that the likelihood of bicycling one
+day is very likely influenced by the choice to bicycle on the days
+immediately preceeding.
+
 %==============================================================================
@@
 
-\subsection{References}
+% \subsection{References}
 
-\begin{thebibliography}{99}
+% \begin{thebibliography}{99}
 
-\bibitem{ref1} J.~Doe, Article name, \textit{Phys. Rev. Lett.}
+% \bibitem{ref1} J.~Doe, Article name, \textit{Phys. Rev. Lett.}
 
-\bibitem{ref2} J.~Doe, J. Smith, Other article name, \textit{Phys. Rev. Lett.}
+% \bibitem{ref2} J.~Doe, J. Smith, Other article name, \textit{Phys. Rev. Lett.}
 
-\bibitem{web} \url{http://www.google.pl}
+% \bibitem{web} \url{http://www.google.pl}
 
-\end{thebibliography}
+% \end{thebibliography}
 %--End of references-----------------------------------------------------------

### 2014-11-18 R/nbinom.R
AVP, AIC, BIC

@@ -101,20 +101,38 @@ mod1_f <- count ~ daylighth + holiday + uw +
     temperatureMax + precipIntensityMax + X +
     Sat + Mon + Tues + Wed + Thurs + Fri # Sunday is reference cat
 
-mod1_m <- glm.nb(mod1_f, data=wb)       # estimate the model
+mod1_m_pois <- glm(mod1_f, data=wb, family=poisson) # poisson for fun
+mod1_m <- glm.nb(mod1_f, data=wb)       # estimate the real model
 set.seed(123456)                        # To reproduce results
 mod1_sb <- simbetas(mod1_m)             # Simulate our betas
 
 
 ## Model fit ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
-## Actual versus predicted plot
+## Actual versus predicted plot - negative binomial
 avpp1 <- ggplot(data.frame(actual=wb$count, predicted=getyhat(mod1_m)),
                 aes(x=actual, y=predicted)) +
          geom_abline(intercept=0, slope=1, color="red") +
          geom_point() + theme_bw()
 
 ggsave(file="avpp1.pdf", path="fig", width=7, height=7, units=("in"))
+
+
+## Actual vs predicted plot - poisson
+avpp2 <- ggplot(data.frame(actual=wb$count, predicted=getyhat(mod1_m_pois)),
+                aes(x=actual, y=predicted)) +
+         geom_abline(intercept=0, slope=1, color="red") +
+         geom_point() + theme_bw()
+
+ggsave(file="avpp2.pdf", path="fig", width=7, height=7, units=("in"))
+
+## AIC
+AIC(mod1_m)                             # Negative Binomial
+AIC(mod1_m_pois)                        # Poisson
+
+## BIC
+BIC(mod1_m)                             # Negative Binomial
+BIC(mod1_m_pois)                        # Poisson
 
 
 ## Counterfactual simulations of model 1 ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

### 2014-11-18 poster/poster.tex
draft poster as printed

@@
 These data are summarized graphically in figures \ref{fg:hist} and
-\ref{fg:timeseries} below. From the histogram, we see what looks to be
-a Poisson distribution with over dispersion. From the timeseries plot,
-we see that there is a clear yearly pattern, as well as a weekly
-pattern.
+\ref{fg:timeseries} below. From the histogram, we see what
+approximately looks to be a Poisson distribution with over dispersion.
+From the timeseries plot, we see that there is a clear yearly pattern,
+as well as a weekly pattern.
 
@@
 \section{Model Specification}
-Because over dispersion was evident in the counts, we chose to fit a
-negative binomial model. This model includes the following variables,
-and was fit using the \texttt{pscl::glm.nb()} function in R.
+Because over dispersion was evident in descriptive exploration of the
+counts, we chose to fit a negative binomial model. This model includes
+the following variables, and was fit using the \texttt{pscl::glm.nb()}
+function in R.
 
@@
 
+\newpage
 \section{Model Fit and Diagnostics}
@@
 from the $y=x$ line, perhaps indicating some other unaccounted source
-of variation in the dataset.
+of variation in the dataset. For comparison, we also fit this model as
+a Poisson model, however both the AIC and BIC were over six times as
+large as in the negative binomial, indicating a much poorer fit.
 
@@
 Similiarly, we see the strong negative effect of precipitation in our simulated
-results.
+results. This corroborates our understanding of bicyclist aversion to rain.
 

### 2014-12-02 R/nbinom.R
Added binned AvP plot, output multiple sizes, add nb llk

@@ -1,5 +1,6 @@
-## install.packages(c("ggplot2", "plyr", "reshape2", "lubridate", "dplyr", "MASS", "pscl"))
-## install_github("pschmied/tile-simcf/simcf") # Requires devtools loaded
+## install.packages(c("ggplot2", "plyr", "reshape2", "lubridate",
+## "dplyr", "MASS", "pscl"))
+## install_github("pschmied/tile-simcf/simcf") # Requires devtools
 library(ggplot2)
 library(plyr)
 library(reshape2)
@@ -67,27 +68,48 @@ getyhat <- function(model) {
 }
 
 binys <- function(y, yhat, by=0.05) {
-    # Bins y according to yhat into bins of some desired size
+    ## Bins y according to yhat into bins of some desired size
     binned <- group_by(
         data.frame(y=y,
                    yhat=yhat,
-                   bin=cut(yhat, seq(0,1,by=by))),
+                   bin=cut(yhat, seq(min(y),max(y),by=by))),
         bin)
     summarize(binned, y=mean(y), yhat=mean(yhat), n=n())
+}
+
+mknb_llk <- function(formula, data) {
+    ## Takes a formula specifying the model and a dataframe to extract
+    ## variables from. Returns a closure to compute the log-likelihood
+    ## of a negative binomial based on the supplied data
+    
+    av <- all.vars(formula)                 # Get variables to be used
+    sdf <- na.omit(subset(data, select=av)) # subset our data, omit NA
+    ys <- sdf[,1]                           # extract dependent var
+    xmat <- cbind(1, as.matrix(sdf[,-1]))   # creat design mat
+
+    function(betaalpha) {
+        expxb <- exp(xmat %*% head(betaalpha, length(betaalpha) - 1))
+        alpha <- tail(betaalpha, 1)
+        theta <- 1/alpha
+        sum(dnbinom(ys, mu=expxb, size=theta, log=TRUE))
+    }
 }
 
 ##
 ## ~~~~~~~~~~~~~~~~~~~~~~~~~~~ Descriptives ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 ## 
 
[…]

### 2014-12-02 R/nbinom.R
change size of avp plot

@@ -156,8 +156,8 @@ avpp2 <- ggplot(avp_bin_m1, aes(x=yhat, y=y, size=n)) +
          xlab("predicted") + ylab("actual") +
          scale_size_continuous(name="# values")
 
-ggsave(file="avpp2.pdf", path="fig", width=7, height=7, units=("in"))
-ggsave(file="avpp2-sm.pdf", path="fig", width=3.5, height=3.5, units=("in"))
+ggsave(file="avpp2.pdf", path="fig", width=7, height=5, units=("in"))
+ggsave(file="avpp2-sm.pdf", path="fig", width=5, height=3.5, units=("in"))
 
 ## Actual vs predicted plot - poisson
 avpp3 <- ggplot(data.frame(actual=wb$count, predicted=getyhat(mod1_m_pois)),

### 2014-12-02 article/article.tex
add initial incomplete article draft; add latex bits to gitignore

% !TeX TS-program = pdflatexmk
% !BIB TS-program = biber

\documentclass[12pt,letterpaper,article]{memoir} % for a short document
\chapterstyle{article}

% Bibliography
\usepackage[authordate,strict,bibencoding=inputenc,doi=false,isbn=false,annotation=true]{biblatex-chicago}

% Graphics
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\addbibresource{../bibliography/bikecounts.bib}

\title{Estimating Bicycle Counts from Seasonal and Weather Factors}
\author{Peter Schmiedeskamp, Weiran Zhao}
%\date{8 October 2014} % Delete this line to display the current date

%%% BEGIN DOCUMENT
\begin{document}

\maketitle

\section*{Introduction}
%
% 1. Explain what we set out to do (i.e. understand influence of
% seasonal and weather factors on bicycling) 2. explain relevance to
% planners (e.g. understanding how much seasonal variability to expect
% for service provision, understand methodologically how to use data
% collected from counters) 3. ??? other things I'm forgetting right
% now.
%

\section*{Literature review}
% 
% Identify where people have talked about weather and seasonal factors
% influencing counts previously. Describe their methods together with any
% limitations, as well as their substantive findings---especially what
% variables they considered. Normally should conduct this first, but
% our driver here is a class assignment! We should fill this in and
% adjust our analysis accordingly before attempting to go to press.
%

\section*{Methodology}
In order to discern the relationship between bicycle counts and
several identified weather, seasonal, and temporal factors, we
developed a statistical model that attempts to predict daily bicycle
counts from these other factors. This section describes the methods
and procedures we used to collect and process the estimation dataset,
the rationale for our selection of variables, our chosen model type,
and our model estimation procedures.
\subsection*{Data collection, processing, description}
%
% Data collection: study site, descriptives of sample, collection from
% forecast.io and seattle.gov, not in that order
%

The data that we attempt to explain were collected at Seattle's
Fremont Bridge and cover a period of two years spanning from October
31, 2012 to October 30, 2014. The Fremont Bridge captures a
substantial number of bicyclists due to its status as one of only five
facilities that carry bicyclists across the canal that separates the
northern and southern halves of Seattle. Bicycle counts are collected
at this location continuously by the City of Seattle using an
in-sidewalk counter manufactured by EcoCounter. When a bicycle passes
over an induction loop embedded in the sidewalk on either side of the
Fremont Bridge, the counter registers the bicycle. Bicyclists may
legally choose to ride in the roadway instead of the sidewalk, and
would not be detected by the counter. However, we believe these
crossings are rare at this location due to the design of the facility,
which directs bicyclists to enter the sidewalk, and from our own
experience riding and observing other riders. Counts are aggregated
into 15 minute intervals, and are made available to the public via the
City of Seattle's data
portal \parencite{City-of-Seattle:aa,City-of-Seattle:ab}.

Weather data are aggregated and made available through Forecast.io's
web services API \parencite{The-Dark-Sky-Company:aa}. Historical daily
summaries are available for a range of weather variables including
several specifically important to our model such as precipitation,
daily minimum and maximum temperatures, sunrise, and sunset.

We downloaded and processed these data programmatically using the R
programming language along with several add-on
packages \parencite{Grolemund:2011aa,Wickham:2011aa,Couture-Beil:2014aa,Lang:2014aa,R-Core-Team:2014aa}.
Bicycle counts were aggregated by day, and then joined to weather data
by date. In addition to the variables collected from these two
sources, we were also interested in controlling for holidays and
whether or not the nearby University of Washington was in session.
These data were collected and coded manually from the University of
Washington's historic academic calendars.

Figures \ref{fg:hist} and \ref{fg:timeseries} provide visual summaries
of the processed counts data.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.9\textwidth]{../fig/dp1}
  \caption{Histogram showing frequency of observed bicycles}
  \label{fg:hist}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.9\textwidth]{../fig/dp2}
  \caption{Timeseries plot of bicycle counts}
  \label{fg:timeseries}
\end{figure}

\subsection*{Variable selection}
% 
% Data analysis: specification of model, choice of model (nb v pois),
% description of counterfactual simulation using the estimated model,
% software used.
%

Out of the set independent variables retrieved from Forecast.io and
the University of Washington, we selected a subset that we felt best
reflected our specific research questions. The final set of variables
included in our model include Table \ref{tb:variables} summarizes the
variables used in our model.

\begin{centering}
\begin{table}[h]
\caption{Variables included in model specification}
\begin{tabular}{ll}
\toprule
Variable & Description \\
\midrule
Count* & Number of bicycles per day \\
Daylight hours & Length of daylight in hours \\
UW in session & Was the University of Washington in session? \\
Holiday & Was the day a holiday as recognized by UW? \\
Max temp & Maximum temperature for the day \\
Max precip & Daily maximum inches of precipitation in any hour \\
Sat--Fri & Day of the week dummy variable (relative to Sunday) \\
Day number & Sequentially numbered day of study \\
\bottomrule
\multicolumn{2}{r}{* Dependent variable}
\end{tabular}

\label{tb:variables}
\end{table}
\end{centering}

Daylight hours (defined as $sunrise - sunset$) and the University of
Washington in-session status were selected to represent seasonality.
Daylight hours was chosen instead of a calendar-based categorization
of season in part because Seattle's Pacific Maritime Climate differs
substantially from traditional notions of Spring, Summer, Autumn, and
Winter. Daylight hours also is measured as a continuous value at a
finer temporal resolution of one day. Finally, daylight hours adjusts
according to latitude, which may make this model estimation procedure
and specification more transferable to other sites in the future.

We deemed the University of Washington variable important in part
because of the Fremont Bridge's proximity and connection via the Burke
Gilman Trail to the University of Washington. We also felt that this
variable was a suitable proxy for the ``school season,'' which more
broadly captures whether or not local schools are in session. The
academic calendars of the various local schools do not align
perfectly, however they still overlap substantially with the
University of Washington, which is itself the largest educational
institution in the region.

Inclusion of the holiday variable was an attempt to account for
some low outlier counts. Upon inspection of the dataset, Christmas and
Thanksgiving in particular had very low counts of bicycles relative to
the days preceding and following. Relatedly, but not accounted for by
any variable in our model, are some of the high outlier counts. Upon
inspection, some of the highest counts were observed on National Bike
to Work Day and on the day of the Fremont Solstice Parade, which
typically draws large numbers of bicyclists as participants and
spectators. The omission of such a variable is justified based on the
few occurrences of high outlier counts, and our desire for this model
to only include variables that could be collected or straightforwardly
adapted to other locations.

Daily maximum temperature, measured in Fahrenheit, was chosen to
represent temperature (rather than, for example, substituting or
adding daily minimum temperature) in part to retain simplicity in the
model, in part because there is relatively little daily temperature
variation in Seattle due to the moderating effect of large water
bodies, and in part because maximum temperature better reflects the
conditions during daylight hours when most bicycle trips would occur.
This simplification may not be warranted for other locations that
experience greater temperature variation than Seattle.

Maximum precipitation, which measures the maximum inches of
precipitation that occurred in any hour throughout the day, was chosen
rather than average precipitation based on the notion that bicyclists
might make travel decisions based on a likely worst case scenario.
This assumption is slightly more problematic than our assumptions
about temperature, in that we do expect bicyclists to be at least
somewhat sensitive to average conditions or conditions observed at
their time of departure. As in the case of temperature, this
simplification would be less justifiable in locations that experience
greater daily variation in precipitation or in locations that have a
predictable pattern of precipitation during certain hours.

Day of the week was added due to its presence in the literature, as
well as an apparent weekly pattern revealed visually by zooming into
the timeseries plot. These data were coded as a set of Boolean dummy
variables, excluding Sunday as the reference category.

The final variable, the day number, was included so that we could test
for a linear trend in bicycling volumes. We created this variable by
sequentially numbering (1--720) the observed counts by day during the study period.

\subsection*{Model estimation}
% Model choice
A natural model choice for count data is the Poisson model, however we
believe that our count data may be overdispersed. Overdispersion, or
contagion between events, violates the mean-variance equivalence
assumption of the Poisson model. While overdispersion would not impact
our parameter estimates, it would result in overly optimistic margins
of error. In order to account for and estimate the amount of
overdispersion present in our data, we chose a negative binomial model
type. Because there were no days observed with zero bicycle counts, we
did not need to resort to a zero-inflated model as is often necessary
when dealing with count data.

% Estimation method
We fit the model in R using the \texttt{glm.nb} function from the MASS
package \parencite{Venables:2002aa}. For comparison, we also estimated a
Poisson model with an analogous specification in R using the
\texttt{glm} function. A much lower AIC and BIC in the case of the
negative binomial model confirmed that it was a better fit than the
Poisson. We then tested for overdispersion using the \texttt{odTest} function
from the Pscl package \parencite{Jackman:2014aa}. The highly significant
chi-square test statistic provided strong evidence that overdispersion
was present in the data, further confirming the choice of a negative
binomial.

In addition to testing the model fit relative to its Poisson
analog, we visually assessed the fit of our negative binomial model by
plotting actual versus predicted values as shown in figure
\ref{fg:avp}. This fit appears to be generally good, though very high
count days are predicted less accurately.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.7\textwidth]{../fig/avpp2-sm}
  \caption{Actual versus predicted values as fit by negative binomial model}
  \label{fg:avp}
\end{figure}

% Model interpretation
In order to provide results that are more readily interpretable by
non-statisticians, we used counterfactual simulation to isolate
individual terms from the model that correspond to our research
questions. In so doing, we generated various quantities of interest
including point estimates and confidence intervals, and then plotted
them for visual inspection. Counterfactual simulations were performed
with a modified version of the Simcf R package, and visualized with
ggplot2 \parencite{Adolph:2014aa,Schmiedeskamp:aa,Wickham:2009aa}.
Results of these simulations are presented in the following section.

\section*{Results}
% 
% One subsection for each research question
% 
This section presents the results from the statistical model and
accompanying counterfactual simulations as described in the preceding
section. Each of the main research questions of seasonality, weather,
and general trend are addressed here. In addition, additional added
control variables such as day of the week are presented.

Without exception, each of the coefficients in our model was
statistically significant at the $p < 0.05$. Further, with the
exception of the Saturday coefficient, all coefficients were
significant at the $p < 0.001$ level. The remainder of this section
focuses on presenting the substantial effect of each variable.

\subsection*{Seasonality}
As discussed previously, we considered two variables to address the
question of seasonality: the first being the number of daylight hours,
and the second being whether or not the University of Washington
(proxying more generally for other educational institutions) is in
sesssion.

As seen in figure \ref{fg:seasonality}, we see a substantial increase
in bicycle volumes when the University of Washington is in session.
With all other factors held constant we see that, on days when the
University is in session, there is an average of approximately 367
additional bicycles observed. Similarly, we see an almost linear
increase in bicycles with increased day length.

\begin{figure}[h!]
  \centering
  \includegraphics[width=.8\textwidth]{../fig/m1c4}
  \caption{Effect of daylight hours and University of Washington
    in-session status on bicycle counts, with shaded 95\% confidence
    regions.}
  \label{fg:seasonality}
\end{figure}

\subsection*{Weather}

\begin{figure}[h!]
  \centering
  \includegraphics[width=.8\textwidth]{../fig/m1c2}
  \caption{Effect of precipitation on counts, with shaded 95\%
    confidence region.}
  \label{fg:precipitation}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=.8\textwidth]{../fig/m1c1}
  \caption{Effect of temperature on counts, with shaded 95\%
    confidence region.}
  \label{fg:temperature}
\end{figure}

\subsection*{General trend in bicycle counts}

\begin{figure}[h!]
  \centering
  \includegraphics[width=.8\textwidth]{../fig/m1c5}
  \caption{General trend in bicycling counts, all other factors held
    constant, with 95\% confidence region.}
  \label{fg:trend}
\end{figure}

\subsection*{Day of week variation}

\begin{figure}[h!]
  \centering
  \includegraphics[width=.8\textwidth]{../fig/m1c3}
  \caption{Variation in counts throughout the week, with 95\% confidence bars.}
  \label{fg:dayofweek}
\end{figure}

\clearpage
\section*{Discussion}

\section*{Conclusion}

\printbibliography
\end{document}

### 2014-12-03 R/nbinom.R
add residuals plots

@@ -159,6 +159,18 @@ avpp2 <- ggplot(avp_bin_m1, aes(x=yhat, y=y, size=n)) +
 ggsave(file="avpp2.pdf", path="fig", width=7, height=5, units=("in"))
 ggsave(file="avpp2-sm.pdf", path="fig", width=5, height=3.5, units=("in"))
 
+## Actual vs predicted plot - nb usiung "predict"
+
+predictedy <- predict(mod1_m,newdata=wb, type="response")
+
+avpp4 <- ggplot(data.frame(actual=wb$count, predictedy),
+                 aes(x=actual, y=predictedy)) +
+       geom_abline(intercept=0, slope=1, color="red") +
+       geom_point() + theme_bw()
+
+ggsave(file="avpp4.pdf", path="fig", width=7, height=5, units=("in"))
+ggsave(file="avpp4-sm.pdf", path="fig", width=5, height=3.5, units=("in"))
+
 ## Actual vs predicted plot - poisson
 avpp3 <- ggplot(data.frame(actual=wb$count, predicted=getyhat(mod1_m_pois)),
                 aes(x=actual, y=predicted)) +
@@ -176,6 +188,14 @@ AIC(mod1_m_pois)                        # Poisson
 BIC(mod1_m)                             # Negative Binomial
 BIC(mod1_m_pois)                        # Poisson
 
+##Influence Plot: Residuals
+residuals <- influencePlot(mod1_m, identify = FALSE)
+residuals_pois <- influencePlot(mod1_m_pois, identify = FALSE)
+
+ggsave(file="residuals.pdf", path="fig", width=7, height=7, units=("in"))
+ggsave(file="residuals.pdf", path="fig", width=3.5, height=3.5, units=("in"))
+ggsave(file="residuals_pois.pdf", path="fig", width=7, height=7, units=("in"))
+ggsave(file="residuals_pois.pdf", path="fig", width=3.5, height=3.5, units=("in"))
 
 ## Counterfactual simulations of model 1 ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 

### 2014-12-07 article/article.tex
add the literature review

@@
 
+Bicycling volumn in cities iuseful for practitioners and researchers to understand safety, travel behavior, and development impacts.
+Therefore it is of great interest to investigate the relationship between bicycle volume and various factors.
+The goal is to build a predictive model based on this relationship, so that one can better estimate and forecast bicycle travel, help bicycle facility planning, etc.
+There have been a lot of papers in the last decade related to this topic (see, for example, \cite{Griswold11,Nosal12, Fields12, Niemeier96}, and reference therein).
+Regarding data source, survey/census data are more used to explain influencing factors such as demographic and socioeconomic factors to mode choice~\cite{Parkin08, Helbich14} while self-collected data(automatic or manual) are used in more studies in recently to track and analyze the count in a long period of time~\cite{Griswold11,Nosal12}.
+Despite a great amount of research, the methodology is more or less the same: 1) First propose a set of explanatory variables; 2) Then fit a proper regression model and justify the result.
+
+According to a recent report by Bassok et.
+al (2011)~\cite{Bassok11}, there are eleven primary indicators, which are time of day~\cite{Schwartz99}, season~\cite{Niemeier96}, population and employment densities~(\cite{Pinjari09,McCahil08}), mix of use~\cite{Pinjari09}, bicycle facility type~\cite{Hunt07}, traffic volume~\cite{McDonald07}, rain and temperature~\cite{Parkin08,Niemeier96}, income~\cite{Turner98}, and age~\cite{Hunt07}.
+This section outlines the key points made in the literature that are relevant to some of most important variables.
+Research has found the variability for counts has a positive association with high temperature and low precipitation~\cite{Parkin08,Niemeier96}.
+Meanwhile, as suggested by Lewin (2011)~\cite{Lewin11} and Thomas (2008)~\cite{Thomas08}, the effects of precipitation and temperature on bicycle volumes are nonlinear.
+For example, bicycle traffic can decrease in both very cold and very hot weather as noted in Richardson (2000)~\cite{Richardson00}.
+Apart from the usual temperature and rain variables, Miranda-Moreno (2011)~\cite{MM11} finds humidity and additional precipitation variables including the presence of rain in the morning and/or during the previous three hours to be significant too.
+Other comparative studies are also available where bicycle counts are conducted in different cities and different sensitives to weather are examined~\cite{Rose11}.
+As for longitudinal study, Niemeier (1996)~\cite{Niemeier96} finds increased variability for counts conducted in the later months of the year.
+Jones et al.
+(2010)~\cite{Jones10} conclude that morning peak hours from 6 AM to 9 PM accounts for a consistent 95\% of the total bicycle volumes by hourly count data.
+The simple linear regression model has been used in the many general applications~\cite{Jones08,Jones10}.
+Few exceptions include Miranda-Moreno (2011)~\cite{MM11} which develops a count model and Thomas (2008)~\cite{Thomas08} which develops a time-series model.
+Niemier (1996) also uses a Poisson model to statistically confirm that many of the factors thought to influence cyclists.
+The work by Gallop et.
+al (12)~\cite{Gallop12} adopts a similar time-series approach while incorporating an autoregressive integrated moving average (ARIMA) analysis.
+
+After reviewing these literature, we found that the bicycle count data is need to be further analysis with a long-term data and a better model.
+How seasonal factors influence bicycle flow needs to be examined in data that last more than a year.
+Plus, among the past researches, almost no literature discuss about the goodness of fit of their modeling,  a model that can better describe and forecast the bicycle count in longitudinal form is necessary to developed.
+Models for count data with better estimation methods are promising.
+
 \section*{Methodology}
@@
 and our model estimation procedures.
+
 \subsection*{Data collection, processing, description}

### 2014-12-08 article/article.tex
Tidy bibliography entries/citations, move to 2 column layout

--- 2014-12-02 article/article.tex
@@
 
-\documentclass[12pt,letterpaper,article]{memoir} % for a short document
+\documentclass[12pt,letterpaper,article,twocolumn]{memoir}
+
+% Layout options
 \chapterstyle{article}
+\setulmarginsandblock{.5in}{.75in}{*}
+\setlrmarginsandblock{.5in}{*}{*}
+\checkandfixthelayout
+\raggedbottom
 
@@
 
-\title{Estimating Bicycle Counts from Seasonal and Weather Factors}
+\title{Estimating Bicycle Counts in Seattle from Seasonal and Weather Factors}
 \author{Peter Schmiedeskamp, Weiran Zhao}
@@
 
+Bicycling offers a variety of social benefits that include the health
+of those choosing to bicycle regularly, as well as more broadly a
+society-wide increase in mobility as people shift from congestion
+causing modes to bicycling. As such, bicycling is being promoted by
+policy makers and urban planners as an increasingly important element
+of urban transportation. In order to develop policy and
+infrastructural improvements to induce more bicycling, a robust
+understanding of factors related to bicycling is required.
+
+This article contributes to that understanding by focusing on the
+substantive effects of weather and seasonality on the number of
+bicyclists observed on a given day. In order to address these two
+primary research questions, we develop a statistical model estimated
+using two years of automated bicycle counts collected on Seattle's
+Fremont Street bridge, as well as historic weather and season data
+collected from other sources.
+
+Using this model, we were able to identify and quantify substantial
+effects on bicycle counts associated with temperature, precipitation,
+hours of daylight, and school session status. Additionally, we
+estimated changes in bicycle numbers associated with the day of the
+week and holidays, as well as an overall trend toward increased
+bicycling.
+
 \section*{Literature review}
@@
 
+Bicycling volumn in cities iuseful for practitioners and researchers
+to understand safety, travel behavior, and development impacts.
+Therefore it is of great interest to investigate the relationship
+between bicycle volume and various factors. The goal is to build a
+predictive model based on this relationship, so that one can better
+estimate and forecast bicycle travel, help bicycle facility planning,
+etc. There have been a lot of papers in the last decade related to
+this topic (see, for example,
+\cite{Griswold:2011aa,Fields:2012aa,Niemeier:1996aa,Nosal:2014aa}, and
+reference therein). Regarding data source, survey/census data are more
+used to explain influencing factors such as demographic and
+socioeconomic factors to mode choice \cite{Parkin:2008aa,Helbich:2014aa} while
+self-collected data(automatic or manual) are used in more studies in
+recently to track and analyze the count in a long period of
+time\cite{Griswold:2011aa,Nosal:2014aa}. Despite a great amount of research,
+the methodology is more or less the same: 1) First propose a set of
+explanatory variables; 2) Then fit a proper regression model and
+justify the result.
+
+According to a recent report by \cite{Bassok:2011aa}, there are eleven
+primary indicators, which are time of day \cite{Schwartz:1999aa},
+season \cite{Niemeier:1996aa}, population and employment densities
+\cite{McCahil:2008aa,Pinjari:2009aa}, mix of use
+\cite{Pinjari:2009aa}, bicycle facility type \cite{Hunt:2007aa},
+traffic volume \cite{McDonald:2007aa}, rain and temperature
+\cite{Niemeier:1996aa,Parkin:2008aa}, income \cite{Turner:1998aa}, and
+age \cite{Hunt:2007aa}. This section outlines the key points made in
+the literature that are relevant to some of most important variables.
+Research has found the variability for counts has a positive
+association with high temperature and low precipitation
+\cite{Niemeier:1996aa,Parkin:2008aa}. Meanwhile, as suggested by
+\cite{Lewin:2011aa} and \cite{Thomas:2009aa}, the effects of
+precipitation and temperature on bicycle volumes are nonlinear. For
+example, bicycle traffic can decrease in both very cold and very hot
+weather as noted in \cite{Richardson:2000aa}. Apart from the usual
+temperature and rain variables, \cite{Miranda-Moreno:2011aa} finds
+humidity and additional precipitation variables including the presence
+of rain in the morning and/or during the previous three hours to be
+significant too. Other comparative studies are also available where
+bicycle counts are conducted in different cities and different
+sensitives to weather are examined \cite{Rose:2011aa}. As for
+longitudinal study, \cite{Niemeier:1996aa} finds increased variability
+for counts conducted in the later months of the year.
+\cite{Jones:2010aa} conclude that morning peak hours from 6 AM to 9 PM
+accounts for a consistent 95\% of the total bicycle volumes by hourly
+count data. The simple linear regression model has been used in the
+many general applications \cite{Jones:2008aa,Jones:2010aa}. Few
+exceptions include \cite{Miranda-Moreno:2011aa} which develops a count
+model and \cite{Thomas:2009aa} which develops a time-series model.
+\cite{Niemeier:1996aa} also uses a Poisson model to statistically
+confirm that many of the factors thought to influence cyclists. The
+work by \cite{Gallop:2012aa} adopts a similar time-series approach
+while incorporating an autoregressive integrated moving average
+(ARIMA) analysis.
+
+After reviewing these literature, we found that the bicycle count data
+is need to be further analysis with a long-term data and a better
+model. How seasonal factors influence bicycle flow needs to be
+examined in data that last more than a year. Plus, among the past
+researches, almost no literature discuss about the goodness of fit of
+their modeling, a model that can better describe and forecast the
+bicycle count in longitudinal form is necessary to developed. Models
+for count data with better estimation methods are promising.
+
 \section*{Methodology}
@@
   \centering
-  \includegraphics[width=0.9\textwidth]{../fig/dp1}
+  \includegraphics[width=0.5\textwidth]{../fig/dp1}
   \caption{Histogram showing frequency of observed bicycles}
@@
   \centering
-  \includegraphics[width=0.9\textwidth]{../fig/dp2}
+  \includegraphics[width=0.5\textwidth]{../fig/dp2}
   \caption{Timeseries plot of bicycle counts}
@@
 \begin{table}[h]
+\begin{scriptsize}
 \caption{Variables included in model specification}
-\begin{tabular}{ll}
+\begin{tabularx}{0.5\textwidth}{>{\raggedright\arraybackslash}p{1.6cm}>{\raggedright\arraybackslash}X}
 \toprule
@@
 Count* & Number of bicycles per day \\
-Daylight hours & Length of daylight in hours \\
-UW in session & Was the University of Washington in session? \\
+Daylight  & Length of daylight in hours \\
+UW & Was the University of Washington in session? \\
 Holiday & Was the day a holiday as recognized by UW? \\
@@
 Sat--Fri & Day of the week dummy variable (relative to Sunday) \\
-Day number & Sequentially numbered day of study \\
+Day \# & Sequentially numbered day of study \\
 \bottomrule
 \multicolumn{2}{r}{* Dependent variable}
-\end{tabular}
+\end{tabularx}
 
 \label{tb:variables}
+\end{scriptsize}
 \end{table}
@@
 
-Daylight hours (defined as $sunrise - sunset$) and the University of
+Daylight hours (defined as $sunset - sunrise$) and the University of
 Washington in-session status were selected to represent seasonality.
@@
 for a linear trend in bicycling volumes. We created this variable by
-sequentially numbering (1--720) the observed counts by day during the study period.
+sequentially numbering (1--720) the observed counts by day during the
+study period.
 
@@
 A natural model choice for count data is the Poisson model, however we
-believe that our count data may be overdispersed. Overdispersion, or
+believed that our count data were overdispersed. Overdispersion, or
 contagion between events, violates the mean-variance equivalence
@@
 type. Because there were no days observed with zero bicycle counts, we
-did not need to resort to a zero-inflated model as is often necessary
+did not need to resort to zero-inflated models as is often necessary
 when dealing with count data.
@@
 \ref{fg:avp}. This fit appears to be generally good, though very high
-count days are predicted less accurately.
+count days are predicted somewhat less accurately.
 
@@
   \centering
-  \includegraphics[width=0.7\textwidth]{../fig/avpp2-sm}
+  \includegraphics[width=0.5\textwidth]{../fig/avpp2-sm}
   \caption{Actual versus predicted values as fit by negative binomial model}
@@
 and general trend are addressed here. In addition, additional added
-control variables such as day of the week are presented.
+control variables such as day of the week, holidays, and linear trend
+are presented.
 
@@
 significant at the $p < 0.001$ level. The remainder of this section
-focuses on presenting the substantial effect of each variable.
+focuses on presenting the substantive effect of each variable.
 
@@
 and the second being whether or not the University of Washington
-(proxying more generally for other educational institutions) is in
-sesssion.
+(proxying more generally for other educational institutions) was in
+session.
 
@@
 With all other factors held constant we see that, on days when the
-University is in session, there is an average of approximately 367
-additional bicycles observed. Similarly, we see an almost linear
+university is in session, there is an average of approximately 367
+additional bicycles observed. Similarly, we see a roughly linear
 increase in bicycles with increased day length.
@@
   \centering
-  \includegraphics[width=.8\textwidth]{../fig/m1c4}
+  \includegraphics[width=.5\textwidth]{../fig/m1c4}
   \caption{Effect of daylight hours and University of Washington
@@
   \centering
-  \includegraphics[width=.8\textwidth]{../fig/m1c2}
+  \includegraphics[width=.5\textwidth]{../fig/m1c2}
   \caption{Effect of precipitation on counts, with shaded 95\%
@@
   \centering
-  \includegraphics[width=.8\textwidth]{../fig/m1c1}
+  \includegraphics[width=.5\textwidth]{../fig/m1c1}
   \caption{Effect of temperature on counts, with shaded 95\%
@@
   \centering
-  \includegraphics[width=.8\textwidth]{../fig/m1c5}
+  \includegraphics[width=.5\textwidth]{../fig/m1c5}
   \caption{General trend in bicycling counts, all other factors held
@@
   \centering
-  \includegraphics[width=.8\textwidth]{../fig/m1c3}
+  \includegraphics[width=.5\textwidth]{../fig/m1c3}
   \caption{Variation in counts throughout the week, with 95\% confidence bars.}

### 2014-12-08 article/article.tex
Finish draft results sections

@@
 
-\title{Estimating Bicycle Counts in Seattle from Seasonal and Weather Factors}
+% Float barriers
+\usepackage[section]{placeins}
+
+\title{Estimating Daily Bicycle Counts in Seattle from Seasonal and Weather Factors}
 \author{Peter Schmiedeskamp, Weiran Zhao}
@@
 
-\subsection*{Model estimation}
+\subsection*{Model estimation and goodness of fit}
 % Model choice
@@
 
-As seen in figure \ref{fg:seasonality}, we see a substantial increase
+Figure \ref{fg:seasonality} shows that we see a substantial increase
 in bicycle volumes when the University of Washington is in session.
@@
 
-\begin{figure}[h!]
+\begin{figure}[htbp!]
   \centering
@@
 
+\FloatBarrier
 \subsection*{Weather}
+As in the case of seasonality, our model represents weather using two
+variables, in this case precipitation (measured as the maximum amount
+of precipitation falling in any hour of that day), and maximum
+temperature (measured daily, in Fahrenheit).
 
-\begin{figure}[h!]
+The effect of precipitation is shown in figure \ref{fg:precipitation}.
+From this, we can see a clear inverse relationship between
+precipitation and bicycle counts. The rate of decrease in bicycles
+appears to begin somewhat steeply, and then begins to slow slightly at
+higher amounts of precipitation. This suggests that people are
+generally more sensitive to the presence of precipitation than the
+intensity.
+
+Temperature, in contrast to precipitation, has a clearly positive
+association with increased numbers of bicyclists. We were somewhat
+surprised to not see a leveling off in counts at very high
+temperatures---likely due to Seattle's moderate summer climate, and the
+lack of extremely warm days in our dataset.
+
+\begin{figure}[ht!]
   \centering
@@
 
-\begin{figure}[h!]
+\begin{figure}[ht!]
   \centering
@@
 
+\FloatBarrier
 \subsection*{General trend in bicycle counts}
+Our results (shown in figure \ref{fg:trend}) confirm the presence of a
+general trend toward increased numbers of bicycles at this location.
+With all else held constant, we would expect to see roughly 300 more
+bicycles on days at the end of our study period than at the beginning.
+This is consistent with figures reported elsewhere that suggest that
+bicycling is increasing in a number of cities including
+Seattle \parencite{League-of-American-Bicyclists:aa}.
 
-\begin{figure}[h!]
+\begin{figure}[ht!]
   \centering
@@
 
+\FloatBarrier
 \subsection*{Day of week variation}
+As discussed previously, day of the week was included due to the
+weekly variation in bicycle counts apparent in timeseries plots.
+Figure \ref{fg:dayofweek} shows several interesting aspects of these
+results. First, we see much higher numbers of bicyclists on weekdays
+than on weekends. This strongly suggests that the majority of the
+bicycle traffic at this location is for commuter purposes.
 
-\begin{figure}[h!]
+Comparing between weekday results, we see that most days are roughly
+the same, with the some drop-off toward Thursday and Friday. This
+decrease toward the end of the work week might be attributable to
+individuals either working non-traditional schedules or perhaps people
+adjusting their travel mode choice in order to accommodate social
+engagements.
+
+\begin{figure}[ht!]
   \centering

### 2014-12-08 article/article.tex
discussion and conclusions draft

@@
 
-\clearpage
+\FloatBarrier
 \section*{Discussion}
+This research set out to better understand the relationship between
+bicycle counts and weather and season. However, weather and season are
+themselves comprised by a number of constituent elements. We believe
+this research has chosen defensible representations of weather and
+season. However, we readily concede that there exist other reasonable
+representations of season and weather.
+
+In particular, others might be more interested in building a model
+that places more emphasis on predictive power. Those individuals might
+choose to add a number of additional variables or consider more
+complex interactions between terms. By contrast, in this model
+formulation, we present a relatively simple model with comparatively
+fewer terms. This was due to our interest in a conceptual
+understanding of the influence of weather and season.
+
+Even within our model, as noted in the methodology section, we
+considered alternative specifications, especially those containing
+different measures of temperature and precipitation. As an example, in
+choosing to represent temperature as the daily maximum, we make some
+assumptions about people and their mode choice decision making
+processes. People cannot know the daily maximum temperature in advance
+of their morning commute, however the measured daily maximum
+temperature is likely to be similar to temperatures reported in
+temperature forecasts. A superior measure might be a consensus of
+forecasts available in the morning of each day. Similarly, if
+Seattle's daily high and daily low temperatures did not have such low
+variation, a daily maximum temperature might be less important to a
+winter-time bicyclist than daily minimum.
+
+One variable missing from our analysis that we would have liked to
+include is cloud cover.\footnote{We gratefully acknowledge Christopher
+  Adolph for this suggestion.} This variable would have allowed us to
+account for the possibility of brilliant, sunny, but cold winter days,
+as well as warm, dry, but gloomy summer days. Unfortunately, the
+dataset downloaded from Forecast.io included a great deal of missing
+values for this variable, and we were unable to identify another
+suitable source for this data in time for this article.
+
+Another limitation of this research comes from our sample of bicycle
+counts. While we do have continuous counts spanning a full two years,
+these counts were taken at just one point location. We believe the
+Fremont bridge is somewhat representative of a high-volume bicycle
+facility in Seattle, however we might, for example, see a decreased
+effect size for University of Washington session status if we looked
+at counts further from the university. The City of Seattle has begun
+collecting and releasing data from other counters, which could be
+included in a future version of this analysis.
+
+Finally, as discussed in the methodology section, we believe the
+choice of the negative binomial model to be a better choice than the
+Poisson, which had been used in some previous studies. A limitation in
+this model, however, is in not accounting for timeseries
+autocorrelation. We think it highly likely that the decision to
+bicycle on one day might influence the decision to bicycle on the
+next. A future direction would be to consider models, such as ARIMA,
+that more explicitly account for time.
 
 \section*{Conclusion}
+This research set out to help clarify the relationship between weather
+and seasonal factors and bicycle counts in Seattle. In order to
+achieve this, we developed a negative binomial model to predict
+bicycle counts based on temperature, precipitation, day length,
+university in-session status, day of the week, and a general linear
+time trend.
+
+For each term included in our model, we found statistically
+significant effects. More importantly, through the use of
+counterfactual simulation, we estimated what we deem to be substantial
+effect sizes associated with each predictor variable.
+
+This article contributes to the existing literature by demonstrating
+the use of a negative binomial model for bicycle counts. In addition,
+the results presented here were generated from data collected over a
+long study period of two years.
+
+While control of the weather and seasons are admittedly beyond the
+scope of policy makers, this research does suggest that planners and
+policy makers may want to develop strategies that help mitigate the
+impacts of the natural environment during the winter months on the
+increasing number of bicyclists. Future research could focus on
+determining what, if any, programmatic or built interventions could
+ameliorate unfavorable cold- and wet-weather bicycling conditions.
 

### 2014-12-08 article/article.tex
make hypotheses more explicit in the introduction

@@
 substantive effects of weather and seasonality on the number of
-bicyclists observed on a given day. In order to address these two
+bicyclists observed on a given day. Specifically, we hypothesize that
+factors associated with inclement weather (low temperature and
+precipitation) will be inversely correlated with bicycle counts. We
+similarly expect to see an additional decrease in counts associated
+with physical and social / institutional aspects of season. That is, in seasons
+where the days get shorter, we expect fewer cyclists due to a kind of
+decision making inertia. We also expect fewer cyclists outside
+the ``school season.''
+
+In order to address these two
 primary research questions, we develop a statistical model estimated

### 2014-12-09 article/article.tex
lit summary table and edits

@@
 
-Bicycling volumn in cities iuseful for practitioners and researchers
+Bicycling volume in cities is useful for practitioners and researchers
 to understand safety, travel behavior, and development impacts.
-Therefore it is of great interest to investigate the relationship
-between bicycle volume and various factors. The goal is to build a
-predictive model based on this relationship, so that one can better
-estimate and forecast bicycle travel, help bicycle facility planning,
-etc. There have been a lot of papers in the last decade related to
-this topic (see, for example,
-\cite{Griswold:2011aa,Fields:2012aa,Niemeier:1996aa,Nosal:2014aa}, and
-reference therein). Regarding data source, survey/census data are more
-used to explain influencing factors such as demographic and
-socioeconomic factors to mode choice \cite{Parkin:2008aa,Helbich:2014aa} while
-self-collected data(automatic or manual) are used in more studies in
-recently to track and analyze the count in a long period of
-time\cite{Griswold:2011aa,Nosal:2014aa}. Despite a great amount of research,
-the methodology is more or less the same: 1) First propose a set of
-explanatory variables; 2) Then fit a proper regression model and
-justify the result.
+Therefore the relationship between bicycle volume and various factors,
+with the goal to build a predictive model based on this relationship,
+has been of great interest to researchers over the last decade (e.g.
+\cite{Griswold:2011aa,Fields:2012aa,Niemeier:1996aa,Nosal:2014aa}).
 
-According to a recent report by \cite{Bassok:2011aa}, there are eleven
-primary indicators, which are time of day \cite{Schwartz:1999aa},
-season \cite{Niemeier:1996aa}, population and employment densities
-\cite{McCahil:2008aa,Pinjari:2009aa}, mix of use
-\cite{Pinjari:2009aa}, bicycle facility type \cite{Hunt:2007aa},
-traffic volume \cite{McDonald:2007aa}, rain and temperature
-\cite{Niemeier:1996aa,Parkin:2008aa}, income \cite{Turner:1998aa}, and
-age \cite{Hunt:2007aa}. This section outlines the key points made in
+To address these questions, researchers have used a variety of
+approaches using different data sources. Studies employing survey or
+census data are more often used to explain influences such as
+demographic and socioeconomic factors on mode choice
+\parencite{Parkin:2008aa,Helbich:2014aa}. In other studies, collected
+count data (from automated counters or manual counts) have more been
+used recently to track and analyze counts over longer periods of
+time \parencite{Griswold:2011aa,Nosal:2014aa}. These counts-based
+approaches have largely followed a similar approach of first proposing
+a set of explanatory variables, fitting some form of regression model,
+and then interpreting / justifying the results vis-\`{a}-vis the
+guiding theory.
+
+A literature review accompanying a recent report by
+\textcite{Bassok:2011aa} identified eleven primary indicators, which are
+time of day \parencite{Schwartz:1999aa}, season \parencite{Niemeier:1996aa},
+population and employment densities
+\parencite{McCahil:2008aa,Pinjari:2009aa}, mix of use
+\parencite{Pinjari:2009aa}, bicycle facility type \parencite{Hunt:2007aa},
+traffic volume \parencite{McDonald:2007aa}, rain and temperature
+\parencite{Niemeier:1996aa,Parkin:2008aa}, income \parencite{Turner:1998aa}, and
+age \parencite{Hunt:2007aa}. This section outlines the key points made in
 the literature that are relevant to some of most important variables.
+
 Research has found the variability for counts has a positive
-association with high temperature and low precipitation
-\cite{Niemeier:1996aa,Parkin:2008aa}. Meanwhile, as suggested by
-\cite{Lewin:2011aa} and \cite{Thomas:2009aa}, the effects of
+association with high temperature and low precipitation. \parencite{Niemeier:1996aa,Parkin:2008aa}. Meanwhile, as suggested by
+\textcite{Lewin:2011aa} and \textcite{Thomas:2009aa}, the effects of
 precipitation and temperature on bicycle volumes are nonlinear. For
 example, bicycle traffic can decrease in both very cold and very hot
-weather as noted in \cite{Richardson:2000aa}. Apart from the usual
-temperature and rain variables, \cite{Miranda-Moreno:2011aa} finds
+weather as noted by \textcite{Richardson:2000aa}. Apart from the usual
+temperature and rain variables, \textcite{Miranda-Moreno:2011aa} finds
 humidity and additional precipitation variables including the presence
@@
 bicycle counts are conducted in different cities and different
-sensitives to weather are examined \cite{Rose:2011aa}. As for
-longitudinal study, \cite{Niemeier:1996aa} finds increased variability
+sensitives to weather are examined \parencite{Rose:2011aa}. As for
+longitudinal studies, \textcite{Niemeier:1996aa} finds increased variability
 for counts conducted in the later months of the year.
-\cite{Jones:2010aa} conclude that morning peak hours from 6 AM to 9 PM
+\textcite{Jones:2010aa} conclude that morning peak hours from 6 AM to 9 PM
 accounts for a consistent 95\% of the total bicycle volumes by hourly
-count data. The simple linear regression model has been used in the
-many general applications \cite{Jones:2008aa,Jones:2010aa}. Few
-exceptions include \cite{Miranda-Moreno:2011aa} which develops a count
+count data. The simple linear regression model has been used in
+several applications \cite{Jones:2008aa,Jones:2010aa}. Other modeling approaches include \cite{Miranda-Moreno:2011aa} which develops a count
 model and \cite{Thomas:2009aa} which develops a time-series model.
 \cite{Niemeier:1996aa} also uses a Poisson model to statistically
-confirm that many of the factors thought to influence cyclists. The
+confirm  many of the factors thought to influence cyclists. The
 work by \cite{Gallop:2012aa} adopts a similar time-series approach
 while incorporating an autoregressive integrated moving average
-(ARIMA) analysis.
+(ARIMA) analysis. A summary of findings from the literature are
+presented in table \ref{tb:lit}.
 
-After reviewing these literature, we found that the bicycle count data
-is need to be further analysis with a long-term data and a better
-model. How seasonal factors influence bicycle flow needs to be
-examined in data that last more than a year. Plus, among the past
-researches, almost no literature discuss about the goodness of fit of
-their modeling, a model that can better describe and forecast the
-bicycle count in longitudinal form is necessary to developed. Models
-for count data with better estimation methods are promising.
+These literature together suggest an opportunity for further model
+development base around long-term automated counts utilizing
+appropriate statistical methods. How seasonal factors influence
+bicycle flow needs to be examined in data that last more than a year.
+One limitation present in much of the past literature is that few
+discuss goodness of fit of their modeling. Further a model that can
+better describe and forecast the bicycle count in longitudinal form is
+necessary to developed. Models for count data with better estimation
+methods offer some promise.
+
+\newfixedcaption{\freetabcaption}{table}
+\begin{centering}
+\begin{scriptsize}
+\freetabcaption{Summary of findings from literature review}
+\begin{ctabular}{p{.6in}p{1.25in}p{1.25in}}
+  \toprule
+  Source & Variable(s) identified & Methods \\
+  \midrule
+  \cite{Fields:2012aa} & Average daily temperature (F); Total weekly precipitation (inches) & Identify patterns through scatter plots; No explicit model is established. \\
+  \cite{Gallop:2012aa} & Temperature, Relative humidity, wind speed, clearness,fog, precipitation & Use ARIMA to account for serial correlation patterns \\
+  \cite{Griswold:2011aa} & Nearby population and employment density, proximity to downtown/freeway, age, education level, income, et.
+& Log linear ordinary least squares regression is used to estimate a bicycle count model \\
+  \cite{Helbich:2014aa} & Daily maiximum air temperature (C), daily average wind speed (m/s), daily precipitation(mm) & Place-specific associations of weather conditions are explored through geographically weighted logit models \\
+  \cite{Hunt:2007aa} & Descriptive variables indicating lane use, secured parking, level of experience, etc. & Logistic model of cycling-related choices \\
+  \cite{Jones:2010aa} & Length of bicycle network, employment density, population density & Standard ordinary least squares regression \\
+  \cite{Lewin:2011aa} & Max temp (F), rain flag, snow flag, weekend flag, over 90 flag & Standard linear regression model \\
+  \cite{McCahil:2008aa} & logarithmic choice measure, population density, worker density & A new space syntax theory is used to evaluate and predict the bicycle volume throughout a network \\
+  \cite{Miranda-Moreno:2011aa} & Temperature (C), humidity (\%), rain presence (0/1), rain in prev.
+3hrs (0/1), warm \& humid (0/1), morning rain (0/1) & Both log-linear model and negative binomial model are tested \\
+  \cite{Niemeier:1996aa} & Morning flag, rain flag, high temp flag, location variable, season variable & A Poisson bicycle count model is assumed and fitted \\
+  \cite{Nosal:2014aa} & Temperature (F), humidity (\%), rain flag, rain prev. 3hrs, am rain, pm rain & The relationship  is analyzed using a log-linear regression model \\
+  \cite{Parkin:2008aa} & Gender, car ownership, hilliness, off-road routes proportion & A logistic regression model of relevant socio-economic and physical variables is estimated. \\
+  \cite{Pinjari:2009aa} & Household density, employment density, fraction of commercial land area, fraction of senior, fraction of race population, etc.
+& The model system takes the form of a joint mixed Multinomial Logit–Multiple Discrete-Continuous Extreme Value (MNL–MDCEV) structure \\
+  \cite{Rose:2011aa} & Temerature(C), rainfall(mm), holiday flag, school seanson flag, day of the week & Weather and other effects examined using an aggregate model of daily ridership \\
+  \cite{Thomas:2009aa} & Temperature(C), sunshine(h), precipitation(mm), wind force(m/s), cycle path use & A bi-level structure is developed with the upper level being a log-linear model and the lower level being a linear model \\
+  \bottomrule
+\label{tb:lit}
+\end{ctabular}
+\end{scriptsize}
+\end{centering}
 
@@
 \begin{centering}
-\begin{table}[h]
+\begin{table}[ht!]
 \begin{scriptsize}

### 2014-12-09 article/article.tex
Readability tweaks to layout, a few assorted edits

@@
 \chapterstyle{article}
-\setulmarginsandblock{.5in}{.75in}{*}
-\setlrmarginsandblock{.5in}{*}{*}
+\setulmarginsandblock{.75in}{1in}{*}
+\setlrmarginsandblock{.75in}{*}{*}
+\setcolsepandrule{.33in}{0pt}
+
 \checkandfixthelayout
@@
 Bicycling offers a variety of social benefits that include the health
-of those choosing to bicycle regularly, as well as more broadly a
-society-wide increase in mobility as people shift from congestion
+of those choosing to bicycle regularly, as well as, more broadly, a
+society-wide increase in mobility when people shift from congestion
 causing modes to bicycling. As such, bicycling is being promoted by
-policy makers and urban planners as an increasingly important element
-of urban transportation. In order to develop policy and
+policy makers and urban planners as an increasingly important
+component of urban transportation. In order to develop policy and
 infrastructural improvements to induce more bicycling, a robust
@@
 bicyclists observed on a given day. Specifically, we hypothesize that
-factors associated with inclement weather (low temperature and
+factors associated with inclement weather (esp. low temperature and
 precipitation) will be inversely correlated with bicycle counts. We
 similarly expect to see an additional decrease in counts associated
-with physical and social / institutional aspects of season. That is, in seasons
-where the days get shorter, we expect fewer cyclists due to a kind of
-decision making inertia. We also expect fewer cyclists outside
-the ``school season.''
+with both physical and social/institutional dimensions of season. That
+is, in seasons where the days get shorter, we expect fewer cyclists
+due to a kind of decision making inertia. We also expect fewer
+cyclists outside the ``school season'' as we believe that a large
+number of bicycling trips are associated with commuting to schools
+that, unlike most other institutions have clear annual cycles.
 
-In order to address these two
-primary research questions, we develop a statistical model estimated
-using two years of automated bicycle counts collected on Seattle's
-Fremont Street bridge, as well as historic weather and season data
-collected from other sources.
+In order to address these two primary research questions, we develop a
+statistical model estimated using two years of automated bicycle
+counts collected on Seattle's Fremont Street bridge, as well as
+historic weather and season data collected from other sources.
 
@@
 traffic volume \parencite{McDonald:2007aa}, rain and temperature
-\parencite{Niemeier:1996aa,Parkin:2008aa}, income \parencite{Turner:1998aa}, and
-age \parencite{Hunt:2007aa}. This section outlines the key points made in
-the literature that are relevant to some of most important variables.
+\parencite{Niemeier:1996aa,Parkin:2008aa},
+income \parencite{Turner:1998aa}, and age \parencite{Hunt:2007aa}.
+This section outlines the key points made in the literature that are
+relevant to some of most important variables.
 
@@
 \freetabcaption{Summary of findings from literature review}
-\begin{ctabular}{p{.6in}p{1.25in}p{1.25in}}
+\begin{ctabular}{p{.6in}p{1.15in}p{1.15in}}
   \toprule
@@
 \caption{Variables included in model specification}
-\begin{tabularx}{0.5\textwidth}{>{\raggedright\arraybackslash}p{1.6cm}>{\raggedright\arraybackslash}X}
+\begin{tabularx}{0.47\textwidth}{>{\raggedright\arraybackslash}p{1.6cm}>{\raggedright\arraybackslash}X}
 \toprule
@@
 Daylight  & Length of daylight in hours \\
-UW & Was the University of Washington in session? \\
-Holiday & Was the day a holiday as recognized by UW? \\
-Max temp & Maximum temperature for the day \\
+UW & The University of Washington in session (TRUE/FALSE) \\
+Holiday & The day a holiday as recognized by UW (TRUE/FALSE) \\
+Max temp & Maximum temperature in Fahrenheit for the day \\
 Max precip & Daily maximum inches of precipitation in any hour \\
-Sat--Fri & Day of the week dummy variable (relative to Sunday) \\
+Sat--Fri & Day of the week dummy variable (TRUE/FALSE, relative to Sunday) \\
 Day \# & Sequentially numbered day of study \\
@@
 impacts of the natural environment during the winter months on the
-increasing number of bicyclists. Future research could focus on
-determining what, if any, programmatic or built interventions could
-ameliorate unfavorable cold- and wet-weather bicycling conditions.
+increasing number of bicyclists. In other words, the delta between
+warm dry days and cold wet days should be treated as the opportunity
+frontier. Future research could focus on determining what, if any,
+programmatic or built interventions could ameliorate unfavorable cold-
+and wet-weather bicycling conditions.
 

### 2014-12-25 R/nbinom.R
influence plots

@@ -95,6 +95,32 @@ mknb_llk <- function(formula, data) {
     }
 }
 
+## Influence plots. This function adapted from:
+## https://gist.github.com/DoktorMike/6278065
+plotInfluence <- function (model, size=10) {
+  if(!inherits(model, "lm")) 
+    stop("You need to supply an lm object.")
+  df<-data.frame(Residual=rstudent(model), 
+                 Leverage=hatvalues(model), 
+                 Cooks=cooks.distance(model), 
+                 Observation=names(hatvalues(model)), 
+                 stringsAsFactors=FALSE)
+  myxint<-c(2*mean(df$Leverage), 3*mean(df$Leverage))
+  inds<-intersect(which(abs(df$Residual) < 2), 
+                  which( df$Leverage < myxint[1]))
+  if(length(inds) > 0) df$Observation[inds]<-""
+  ggplot(df, aes_string(x='Leverage', y='Residual', 
+                        size='Cooks', label='Observation'), 
+         legend=FALSE) +
+    geom_point(colour="black", alpha=0.2) + 
+    scale_size_area(max_size=size) + 
+    theme_bw() + 
+    geom_hline(yintercept=c(2,-2), linetype="dashed") + 
+    geom_vline(xintercept=myxint, linetype="dashed") + 
+    ylab("Studentized Residuals") + 
+    xlab("Hat-Values") + labs(size="Cook's distance")
+}
+
 ##
 ## ~~~~~~~~~~~~~~~~~~~~~~~~~~~ Descriptives ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 ## 
@@ -189,13 +215,13 @@ BIC(mod1_m)                             # Negative Binomial
 BIC(mod1_m_pois)                        # Poisson
 
 ##Influence Plot: Residuals
-residuals <- influencePlot(mod1_m, identify = FALSE)
-residuals_pois <- influencePlot(mod1_m_pois, identify = FALSE)
+residuals <- plotInfluence(mod1_m) + scale_y_continuous(limits=c(-50, 40))
+ggsave(file="residuals.pdf", path="fig", width=7, height=3.5, units=("in"))
+ggsave(file="residuals-small.pdf", path="fig", width=3.5, height=3.5, units=("in"))
 
-ggsave(file="residuals.pdf", path="fig", width=7, height=7, units=("in"))
[…]

### 2014-12-25 article/article.tex
final rev for class

@@
 A literature review accompanying a recent report by
-\textcite{Bassok:2011aa} identified eleven primary indicators, which are
+\textcite{Bassok:2011aa} identified eleven primary indicators. These included 
 time of day \parencite{Schwartz:1999aa}, season \parencite{Niemeier:1996aa},
 population and employment densities
-\parencite{McCahil:2008aa,Pinjari:2009aa}, mix of use
+\parencite{McCahil:2008aa,Pinjari:2009aa}, land-use mix
 \parencite{Pinjari:2009aa}, bicycle facility type \parencite{Hunt:2007aa},
@@
 significant too. Other comparative studies are also available where
-bicycle counts are conducted in different cities and different
+bicycle counts are conducted in different cities, and different
 sensitives to weather are examined \parencite{Rose:2011aa}. As for
@@
 count data. The simple linear regression model has been used in
-several applications \cite{Jones:2008aa,Jones:2010aa}. Other modeling approaches include \cite{Miranda-Moreno:2011aa} which develops a count
-model and \cite{Thomas:2009aa} which develops a time-series model.
-\cite{Niemeier:1996aa} also uses a Poisson model to statistically
+several applications \parencite{Jones:2008aa,Jones:2010aa}. Other modeling approaches include \textcite{Miranda-Moreno:2011aa} which develops a count
+model and \textcite{Thomas:2009aa} which develops a time-series model.
+\textcite{Niemeier:1996aa} also uses a Poisson model to statistically
 confirm  many of the factors thought to influence cyclists. The
-work by \cite{Gallop:2012aa} adopts a similar time-series approach
+work by \textcite{Gallop:2012aa} adopts a similar time-series approach
 while incorporating an autoregressive integrated moving average
@@
   \cite{Pinjari:2009aa} & Household density, employment density, fraction of commercial land area, fraction of senior, fraction of race population, etc.
-& The model system takes the form of a joint mixed Multinomial Logit–Multiple Discrete-Continuous Extreme Value (MNL–MDCEV) structure \\
+& The model system takes the form of a joint mixed Multinomial Logit–Multiple Discrete-Continuous Extreme Value (MNL-MDCEV) structure \\
   \cite{Rose:2011aa} & Temerature(C), rainfall(mm), holiday flag, school seanson flag, day of the week & Weather and other effects examined using an aggregate model of daily ridership \\
@@
 
+\FloatBarrier
 \subsection*{Data collection, processing, description}
@@
 31, 2012 to October 30, 2014. The Fremont Bridge captures a
-substantial number of bicyclists due to its status as one of only five
-facilities that carry bicyclists across the canal that separates the
-northern and southern halves of Seattle. Bicycle counts are collected
-at this location continuously by the City of Seattle using an
-in-sidewalk counter manufactured by EcoCounter. When a bicycle passes
-over an induction loop embedded in the sidewalk on either side of the
-Fremont Bridge, the counter registers the bicycle. Bicyclists may
-legally choose to ride in the roadway instead of the sidewalk, and
-would not be detected by the counter. However, we believe these
-crossings are rare at this location due to the design of the facility,
-which directs bicyclists to enter the sidewalk, and from our own
-experience riding and observing other riders. Counts are aggregated
-into 15 minute intervals, and are made available to the public via the
-City of Seattle's data
+substantial amount of bicycle traffic due to its status as one of only
+five facilities that carry bicyclists across the canal separating the
+northern and southern halves of Seattle.
+
+Bicycle counts were collected at this location continuously by the
+City of Seattle using an in-sidewalk counter manufactured by
+EcoCounter. When a bicycle passes over an induction loop embedded in
+the sidewalk on either side of the Fremont Bridge, the counter
+registers the bicycle. Bicyclists may legally choose to ride in the
+roadway instead of the sidewalk, and would thus not be detected by the
+counter. However, we believe these crossings are rare at this location
+due to the design of the facility, which directs bicyclists to enter
+the sidewalk, and from our own experience riding and observing other
+riders. Counts are aggregated into 15 minute intervals by the City of
+Seattle, and are made available to the public via the City of
+Seattle's data
 portal \parencite{City-of-Seattle:aa,City-of-Seattle:ab}.
 
-Weather data are aggregated and made available through Forecast.io's
-web services API \parencite{The-Dark-Sky-Company:aa}. Historical daily
+Weather data are collected by a variety of sources and are aggregated
+by Forecast.io. These data are available through the company's web
+services API \parencite{The-Dark-Sky-Company:aa}. Historical daily
 summaries are available for a range of weather variables including
@@
 programming language along with several add-on
-packages \parencite{Grolemund:2011aa,Wickham:2011aa,Couture-Beil:2014aa,Lang:2014aa,R-Core-Team:2014aa}.
+packages \parencite{Grolemund:2011aa,Wickham:2011aa,Couture-Beil:2014aa,Lang:2014aa,R-Core-Team:2014aa}.\footnote{In
+addition, the code used to run the analysis, as well as the final
+dataset are available as a github repository: \url{https://github.com/pschmied/bikecounts-mle}}
 Bicycle counts were aggregated by day, and then joined to weather data
@@
 Figures \ref{fg:hist} and \ref{fg:timeseries} provide visual summaries
-of the processed counts data.
+of the processed counts data. Some apparent outliers are visible at
+the rightmost portion of the histogram. The two highest counts
+occurred the Monday and Tuesday preceding the beginning of National
+Bike to Work Month. And the third highest count occurred on National
+Bike to Work Day.
 
@@
 
+\FloatBarrier
 \subsection*{Variable selection}
@@
 Daylight  & Length of daylight in hours \\
-UW & The University of Washington in session (TRUE/FALSE) \\
-Holiday & The day a holiday as recognized by UW (TRUE/FALSE) \\
+UW & The University of Washington was in session (TRUE/FALSE) \\
+Holiday & The day was a holiday as recognized by UW (TRUE/FALSE) \\
 Max temp & Maximum temperature in Fahrenheit for the day \\
@@
 according to latitude, which may make this model estimation procedure
-and specification more transferable to other sites in the future.
+and specification more transferable to other sites in the future,
+perhaps by interacting latitude with daylight hours.
 
@@
 variable was a suitable proxy for the ``school season,'' which more
-broadly captures whether or not local schools are in session. The
-academic calendars of the various local schools do not align
+broadly captures whether or not other local schools are in session.
+The academic calendars of the various local schools do not align
 perfectly, however they still overlap substantially with the
@@
 
+\FloatBarrier
 \subsection*{Model estimation and goodness of fit}
@@
 We fit the model in R using the \texttt{glm.nb} function from the MASS
-package \parencite{Venables:2002aa}. For comparison, we also estimated a
-Poisson model with an analogous specification in R using the
+package \parencite{Venables:2002aa}. For comparison, we also estimated
+a Poisson model with an analogous specification in R using the
 \texttt{glm} function. A much lower AIC and BIC in the case of the
 negative binomial model confirmed that it was a better fit than the
-Poisson. We then tested for overdispersion using the \texttt{odTest} function
-from the Pscl package \parencite{Jackman:2014aa}. The highly significant
-chi-square test statistic provided strong evidence that overdispersion
-was present in the data, further confirming the choice of a negative
-binomial.
+Poisson. We then tested for overdispersion using the \texttt{odTest}
+function from the Pscl package \parencite{Jackman:2014aa}. The highly
+significant chi-squared test statistic provided strong evidence that
+overdispersion was present in the data.
 
-In addition to testing the model fit relative to its Poisson
-analog, we visually assessed the fit of our negative binomial model by
-plotting actual versus predicted values as shown in figure
-\ref{fg:avp}. This fit appears to be generally good, though very high
-count days are predicted somewhat less accurately.
+Additionally, we compared the residuals plots for both the Poisson and
+negative binomial models. As seen in figures \ref{fg:resnb} and
+\ref{fg:respois}, the residual variance is much lower in the negative
+binomial model, further confirming the choice of a negative
+binomial for these counts.
+
+As a final model comparison, we visually assessed the overall fit of
+our negative binomial model by plotting actual versus predicted values
+as shown in figure \ref{fg:avp}. This fit appears to be generally
+good, though very high count days are predicted somewhat less
+accurately.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{../fig/residuals_pois}
+  \caption{Residual variance from the Poisson model}
+  \label{fg:respois}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.5\textwidth]{../fig/residuals}
+  \caption{Residual variance from the negative binomial model}
+  \label{fg:resnb}
+\end{figure}
 
@@
 individual terms from the model that correspond to our research
-questions. In so doing, we generated various quantities of interest
+questions. In so doing, we simulated various quantities of interest
 including point estimates and confidence intervals, and then plotted
@@
 Without exception, each of the coefficients in our model was
-statistically significant at the $p < 0.05$. Further, with the
+statistically significant at the $p < 0.05$ level. Further, with the
 exception of the Saturday coefficient, all coefficients were
@@
 
+\FloatBarrier
 \subsection*{Seasonality}
@@
 Figure \ref{fg:seasonality} shows that we see a substantial increase
-in bicycle volumes when the University of Washington is in session.
+in bicycle volumes when the University of Washington was in session.
 With all other factors held constant we see that, on days when the
-university is in session, there is an average of approximately 367
-additional bicycles observed. Similarly, we see a roughly linear
+university is in session, we would expect an average of approximately
+367 additional bicycles observed. Similarly, we see a roughly linear
 increase in bicycles with increased day length.

### 2015-07-30 article/article.tex
format for TRB

@@
 
-\documentclass[12pt,letterpaper,article,twocolumn]{memoir}
+\documentclass[12pt,letterpaper,article]{memoir}
 
@@
 \chapterstyle{article}
-\setulmarginsandblock{.75in}{1in}{*}
-\setlrmarginsandblock{.75in}{*}{*}
+\setulmarginsandblock{1in}{1in}{*}
+\setlrmarginsandblock{1in}{*}{*}
 \setcolsepandrule{.33in}{0pt}
@@
 
+% Font
+\usepackage{tgtermes}
+
+% TRB Line numbers
+\usepackage{lineno}
+\usepackage{fnlineno}
+\linenumbers
+
 % Graphics
@@
 \addbibresource{../bibliography/bikecounts.bib}
+
+% Needed for sub-figures
+\newsubfloat{figure}
 
@@
 
+% Header and footer
+\makepagestyle{trb}
+\makeevenhead{trb}{Schmiedeskamp, Zhao}{}{\thepage}
+\makeoddhead{trb}{Schmiedeskamp, Zhao}{}{\thepage}
+
+% Headings styles
+\makechapterstyle{trb}{%
+\renewcommand{\chapterheadstart}{\vspace*{3pc}}
+\renewcommand{\chapnamefont}{\bfseries\raggedright}
+\renewcommand{\chapnumfont}{\bfseries\raggedright}
+\renewcommand{\chaptitlefont}{\bfseries\raggedright}
+\renewcommand{\printchaptername}{\S}
+\renewcommand{\chapternamenum}{ }
+\renewcommand{\printchapternum}{\chapnumfont \thechapter}
+\renewcommand{\afterchapternum}{. }
+\renewcommand{\afterchaptertitle}{\par\nobreak\vspace*{1pc}}
+\renewcommand*{\printchaptertitle}[1]{\chaptitlefont\MakeUppercase{##1}} 
+}
+\setsecheadstyle{\bfseries\raggedright}
+\setsubsecheadstyle{\itshape\raggedright} % danger zone
+
+% Table and figure captions
+% http://tex.stackexchange.com/questions/75654/uppercase-until-end-of-group
+\makeatletter
+\def\redeffloatcapstyle#1{%
+  \expandafter\renewcommand\expandafter{\csname fnum@#1\endcsname}[1]{%
+    \bfseries\MakeUppercase{%
+      \expandafter{\csname #1name\endcsname}~\expandafter{\csname the#1\endcsname}\enskip
+    }
+  }
+}
+\makeatother
+\redeffloatcapstyle{figure}
+\redeffloatcapstyle{table}
+%\captionnamefont{\bfseries\scshape}
+\captiontitlefont{\bfseries}
+\captiondelim{}
+
+\pagestyle{trb}
+\chapterstyle{trb}
+
 %%% BEGIN DOCUMENT
@@
 
-\maketitle
+\thispagestyle{empty}
 
-\section*{Introduction}
+\par\noindent\textbf{ESTIMATING DAILY BICYCLE COUNTS IN SEATTLE FROM SEASONAL AND WEATHER FACTORS}
+
+\vspace*{4pc}
+
+\par\noindent\textbf{Peter Schmiedeskamp} (Corresponding author)\\
+University of Washington, Department of Urban Design and Planning\\
+Box 355740, Seattle, WA 98195-5740\\
+Tel: 206.289.0694\\
+email: pschmied@uw.edu
+
+\vspace*{1.5pc}
+\par\noindent\textbf{Weiran Zhao}\\
+University of Washington, Department of Urban Design and Planning\\
+Box 355740, Seattle, WA 98195-5740\\
+Tel: 206.310.7065\\
+email: weiran36@uw.edu
+
+\vspace*{3pc}
+\par\noindent Word count: \textbf{7,452} (4,952 text \& bibliography + 8 figures + 2 tables)
+
+\vspace*{3pc}
+\par\noindent Submission Date: \textbf{1 August 2015}
+
+\clearpage
+
+\par\noindent\textbf{ABSTRACT}
+
+\par\noindent This paper examines the relationship between several seasonal and
+weather factors and bicycle ridership based on two years of automated
+bicycle counts at a Seattle location. In so doing, we fitted a
+negative binomial model, and then estimated quantities of interest
+using counterfactual simulation. Our findings confirm the significance
+of season (+), temperature (+), precipitation (-), as well as holidays
+(-), day of the week (+ for Mon-Sat, relative to Sun.), and an overall
+trend (+). This paper improves on prior work by demonstrating the use
+of the negative binomial instead of a Poisson model, which is
+appropriate given the potential for overdispersion as observed in
+these data. In addition to validating the significance of factors
+identified from the literature, this paper contributes
+methodologically through its intuitive visualization of effect sizes
+to non-statistical audiences. We believe that the combination of model
+type and counterfactual simulation / visualization presented here
+reflect a reasonable compromise between model complexity and
+intepretability. Results such as these can aid policymakers and
+planners both in understanding bicycle travel demand elasticities, and
+to guide interventions aimed at increasing rates of
+bicycling. Finally, the methods presented here are fully reproducible
+and invite adaptation to other locations.
+
+\vspace*{5pc}
+\par\noindent\textit{Keywords}: Bicycle counts; negative binomial; season; weather; automated; counterfactual simulation; visualization
+
+\clearpage
+
+\chapter*{Introduction}
 %
@@
 similarly expect to see an additional decrease in counts associated
-with both physical and social/institutional dimensions of season. That
-is, in seasons where the days get shorter, we expect fewer cyclists
-due to a kind of decision making inertia. We also expect fewer
-cyclists outside the ``school season'' as we believe that a large
-number of bicycling trips are associated with commuting to schools
-that, unlike most other institutions have clear annual cycles.
+with both physical and social / institutional dimensions of
+season. That is, in seasons where the days get shorter, we expect
+fewer cyclists due to a kind of decision making inertia. We also
+expect fewer cyclists outside the ``school season'' as we believe that
+a large number of bicycling trips are associated with commuting to
+schools that, unlike most other institutions have clear annual cycles.
 
@@
 
-\section*{Literature review}
+\chapter*{Literature Review}
 % 
@@
 (ARIMA) analysis. A summary of findings from the literature are
-presented in table \ref{tb:lit}.
+presented in Table \ref{tb:lit}.
 
@@
 better describe and forecast the bicycle count in longitudinal form is
-necessary to developed. Models for count data with better estimation
-methods offer some promise.
+necessary to be developed. Models for count data with better
+estimation methods offer some promise.
 
-\newfixedcaption{\freetabcaption}{table}
-\begin{centering}
+\begin{table}[htdp]
+\begin{center}
 \begin{scriptsize}
-\freetabcaption{Summary of findings from literature review}
-\begin{ctabular}{p{.6in}p{1.15in}p{1.15in}}
+\caption{Summary of findings from literature review}
+\begin{tabularx}{1.0\textwidth}{>{\raggedright\arraybackslash}p{2.7cm}>{\raggedright\arraybackslash}X>{\raggedright\arraybackslash}X}
   \toprule
@@
   \midrule
-  \cite{Fields:2012aa} & Average daily temperature (F); Total weekly precipitation (inches) & Identify patterns through scatter plots; No explicit model is established. \\
-  \cite{Gallop:2012aa} & Temperature, Relative humidity, wind speed, clearness,fog, precipitation & Use ARIMA to account for serial correlation patterns \\
-  \cite{Griswold:2011aa} & Nearby population and employment density, proximity to downtown/freeway, age, education level, income, et.
+  \cite{Fields:2012aa} & Average daily temperature; Total weekly precipitation & Identify patterns through scatter plots; No explicit model is established. \\
+\\
+  \cite{Gallop:2012aa} & Temperature, Relative humidity, wind speed, visibility, fog, precipitation & Use ARIMA to account for serial correlation patterns \\
+\\
+  \cite{Griswold:2011aa} & Nearby population and employment density, proximity to downtown/freeway, age, education level, income, etc.
 & Log linear ordinary least squares regression is used to estimate a bicycle count model \\
-  \cite{Helbich:2014aa} & Daily maiximum air temperature (C), daily average wind speed (m/s), daily precipitation(mm) & Place-specific associations of weather conditions are explored through geographically weighted logit models \\
+\\
+  \cite{Helbich:2014aa} & Daily maiximum air temperature, daily average wind speed, daily precipitation & Place-specific associations of weather conditions are explored through geographically weighted logit models \\
+\\
   \cite{Hunt:2007aa} & Descriptive variables indicating lane use, secured parking, level of experience, etc. & Logistic model of cycling-related choices \\
+\\
   \cite{Jones:2010aa} & Length of bicycle network, employment density, population density & Standard ordinary least squares regression \\
-  \cite{Lewin:2011aa} & Max temp (F), rain flag, snow flag, weekend flag, over 90 flag & Standard linear regression model \\
+\\
+  \cite{Lewin:2011aa} & Max temp, rain flag, snow flag, weekend flag, over 90 flag & Standard linear regression model \\
+\\
   \cite{McCahil:2008aa} & logarithmic choice measure, population density, worker density & A new space syntax theory is used to evaluate and predict the bicycle volume throughout a network \\
-  \cite{Miranda-Moreno:2011aa} & Temperature (C), humidity (\%), rain presence (0/1), rain in prev.
-3hrs (0/1), warm \& humid (0/1), morning rain (0/1) & Both log-linear model and negative binomial model are tested \\
-  \cite{Niemeier:1996aa} & Morning flag, rain flag, high temp flag, location variable, season variable & A Poisson bicycle count model is assumed and fitted \\
-  \cite{Nosal:2014aa} & Temperature (F), humidity (\%), rain flag, rain prev. 3hrs, am rain, pm rain & The relationship  is analyzed using a log-linear regression model \\
+\\
+  \cite{Miranda-Moreno:2011aa} & Temperature, percent humidity, rain presence, rain presence in prev.
+3hrs, warm \& humid, morning rain & Both log-linear model and negative binomial model are tested \\
+\\
+  \cite{Niemeier:1996aa} & Morning flag, rain flag, high temp flag, location variable, season variable & A Poisson count model is assumed and fitted \\
+\\
+  \cite{Nosal:2014aa} & Temperature, percent humidity, rain flag, rain prev. 3hrs, am rain, pm rain & The relationship is analyzed using a log-linear regression model \\
+\\
   \cite{Parkin:2008aa} & Gender, car ownership, hilliness, off-road routes proportion & A logistic regression model of relevant socio-economic and physical variables is estimated. \\
-  \cite{Pinjari:2009aa} & Household density, employment density, fraction of commercial land area, fraction of senior, fraction of race population, etc.
-& The model system takes the form of a joint mixed Multinomial Logit–Multiple Discrete-Continuous Extreme Value (MNL-MDCEV) structure \\
-  \cite{Rose:2011aa} & Temerature(C), rainfall(mm), holiday flag, school seanson flag, day of the week & Weather and other effects examined using an aggregate model of daily ridership \\
-  \cite{Thomas:2009aa} & Temperature(C), sunshine(h), precipitation(mm), wind force(m/s), cycle path use & A bi-level structure is developed with the upper level being a log-linear model and the lower level being a linear model \\
+\\
+  \cite{Pinjari:2009aa} & Household density, employment density, fraction of commercial land area, demographic factors including proportion of population that are seniors and proportion of population by race & The model system takes the form of a joint mixed Multinomial Logit–Multiple Discrete-Continuous Extreme Value (MNL-MDCEV) structure \\
+\\
+  \cite{Rose:2011aa} & Temperature, rainfall, holiday flag, school season flag, day of the week & Weather and other effects examined using an aggregate model of daily ridership \\
+\\
+  \cite{Thomas:2009aa} & Temperature, sunshine, precipitation, wind force, cycle path use & A bi-level structure is developed with the upper level being a log-linear model and the lower level being a linear model \\
   \bottomrule
 \label{tb:lit}
-\end{ctabular}
+\end{tabularx}
 \end{scriptsize}
-\end{centering}
+\end{center}
+\end{table}
 
-\section*{Methodology}
+\FloatBarrier
+
+\chapter*{Methodology}
+
 In order to discern the relationship between bicycle counts and
@@
 \FloatBarrier
-\subsection*{Data collection, processing, description}
+\section*{Data collection, processing, description}
 %
@@
 
-The data that we attempt to explain were collected at Seattle's
+The data that we attempt to predict were collected at Seattle's
 Fremont Bridge and cover a period of two years spanning from October
@@
 
-Figures \ref{fg:hist} and \ref{fg:timeseries} provide visual summaries
-of the processed counts data. Some apparent outliers are visible at
-the rightmost portion of the histogram. The two highest counts
-occurred the Monday and Tuesday preceding the beginning of National
-Bike to Work Month. And the third highest count occurred on National
-Bike to Work Day.
+Figure \ref{fg:descriptives} provides a visual summary of the processed
+counts data. Some apparent outliers are visible at the rightmost
+portion of the histogram. The two highest counts occurred the Monday
+and Tuesday preceding the beginning of National Bike to Work
+Month. And the third highest count occurred on National Bike to Work
+Day.
 
-\begin{figure}[h!]
+\begin{figure}[ht!]
   \centering
-  \includegraphics[width=0.5\textwidth]{../fig/dp1}
-  \caption{Histogram showing frequency of observed bicycles}
-  \label{fg:hist}
+  \subbottom[Frequency of observed counts\label{fig:top}]%
+    {\includegraphics[width=0.5\textwidth]{../fig/dp1-crop}}\hfill
+  \subbottom[Timeseries of counts\label{fig:bottom}]%
+    {\includegraphics[width=0.5\textwidth]{../fig/dp2-crop}}
+  \caption{Descriptive visualization of bicycle counts dataset.}
+  \label{fg:descriptives}
 \end{figure}
 
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.5\textwidth]{../fig/dp2}
-  \caption{Timeseries plot of bicycle counts}
-  \label{fg:timeseries}
-\end{figure}
+% \begin{figure}[h!]
+%   \centering
+%   \includegraphics[width=0.5\textwidth]{../fig/dp2}
+%   \caption{Timeseries plot of bicycle counts}
+%   \label{fg:timeseries}
+% \end{figure}
 
 \FloatBarrier
-\subsection*{Variable selection}
+\section*{Variable selection}
 % 
@@
 
-Out of the set independent variables retrieved from Forecast.io and
+Out of the set of independent variables retrieved from Forecast.io and
 the University of Washington, we selected a subset that we felt best
 reflected our specific research questions. The final set of variables
-included in our model include Table \ref{tb:variables} summarizes the
-variables used in our model.
+included in our model include those presented and summarized in Table
+\ref{tb:variables}. Further explanation and justification of these
+variable choices follow.
 
-\begin{centering}
 \begin{table}[ht!]
+\begin{center}
 \begin{scriptsize}
 \caption{Variables included in model specification}
-\begin{tabularx}{0.47\textwidth}{>{\raggedright\arraybackslash}p{1.6cm}>{\raggedright\arraybackslash}X}
+\begin{tabularx}{0.7\textwidth}{>{\raggedright\arraybackslash}p{1.6cm}>{\raggedright\arraybackslash}X}
 \toprule
@@
 Max temp & Maximum temperature in Fahrenheit for the day \\
+Max temp\textsuperscript{2} & \texttt{Max temp} squared \\
 Max precip & Daily maximum inches of precipitation in any hour \\
@@
 \end{scriptsize}
+\end{center}
 \end{table}
-\end{centering}
 
@@
 This simplification may not be warranted for other locations that
-experience greater temperature variation than Seattle.
+experience greater temperature variation than Seattle. In addition to
+daily maximum temperature, we also included a daily maximum
+temperature squared term in order to identify the potential for a
+leveling off of counts in very high temperatures.
 
@@
 \FloatBarrier
-\subsection*{Model estimation and goodness of fit}
+\section*{Model estimation and goodness of fit}
 % Model choice
@@
 Additionally, we compared the residuals plots for both the Poisson and
-negative binomial models. As seen in figures \ref{fg:resnb} and
-\ref{fg:respois}, the residual variance is much lower in the negative
-binomial model, further confirming the choice of a negative
-binomial for these counts.
+negative binomial models. As seen in figure \ref{fg:respois}, the
+residual variance is much lower in the negative binomial model,
+further confirming the choice of a negative binomial for these counts.
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[ht!]
   \centering
-  \includegraphics[width=0.5\textwidth]{../fig/residuals_pois}
-  \caption{Residual variance from the Poisson model}
+  \subbottom[Poisson\label{fig:mdleft}]%
+    {\includegraphics[width=0.5\textwidth]{../fig/residuals_pois-crop}}\hfill
+  \subbottom[Negative binomial\label{fig:mdright}]%
+    {\includegraphics[width=0.5\textwidth]{../fig/residuals-crop}}
+  \caption{Residual variance of Poisson vs. negative binomial model.}
   \label{fg:respois}
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{../fig/residuals}
-  \caption{Residual variance from the negative binomial model}
-  \label{fg:resnb}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
   \includegraphics[width=0.5\textwidth]{../fig/avpp2-sm}
-  \caption{Actual versus predicted values as fit by negative binomial model}
+  \caption{Actual versus predicted values as fit by negative binomial model.}
   \label{fg:avp}
@@
 
-\section*{Results}
+\chapter*{Results}
 % 
@@
 
-Without exception, each of the coefficients in our model was
-statistically significant at the $p < 0.05$ level. Further, with the
-exception of the Saturday coefficient, all coefficients were
-significant at the $p < 0.001$ level. The remainder of this section
-focuses on presenting the substantive effect of each variable.
+With the exception of the \texttt{Max temp\textsuperscript{2}} term,
+each of the coefficients in our model was statistically significant at
+the $p < 0.05$ level. Further, with the exception of the \texttt{Max
+  temp\textsuperscript{2}} and \texttt{Saturday} coefficients , all
+coefficients were significant at the $p < 0.001$ level. While this is
+instructive in confirming the significance of these factors, the
+remainder of this section focuses on presenting the substantive effect
+of each variable.
 
 \FloatBarrier
-\subsection*{Seasonality}
+\section*{Seasonality}
 As discussed previously, we considered two variables to address the
@@
 \FloatBarrier
-\subsection*{Weather}
-As in the case of seasonality, our model represents weather using two
-variables, in this case precipitation (measured as the maximum amount
-of precipitation falling in any hour of that day), and maximum
-temperature (measured daily, in Fahrenheit).
+\section*{Weather}
+For our examination of seasonality, our model represents weather using
+three variables. In this case, these are precipitation (measured as
+the maximum amount of precipitation falling in any hour of that day),
+maximum temperature (measured daily, in Fahrenheit), and maximum
+temperature squared.
 
@@
 Temperature, in contrast to precipitation, has a clearly positive
-association with increased numbers of bicyclists. We were somewhat
-surprised to not see a leveling off in counts at very high
-temperatures---likely due to Seattle's moderate summer climate, and the
-lack of extremely warm days in our dataset.
+association with increased numbers of bicyclists. Temperature squared
+was not significant in our model at the $p < 0.05$ level. We were
+somewhat surprised to not see evidence for a leveling off in counts at
+very high temperatures---likely due to Seattle's moderate summer
+climate, and the lack of extremely warm days in our dataset.
 
@@
 \FloatBarrier
-\subsection*{General trend in bicycle counts}
+\section*{General trend in bicycle counts}
 Our results (shown in figure \ref{fg:trend}) confirm the presence of a
@@
 \FloatBarrier
-\subsection*{Day of week variation}
+\section*{Day of week variation}
 As discussed previously, day of the week was included due to the
@@
 \FloatBarrier
-\section*{Discussion}
+\chapter*{Discussion}
 This research set out to better understand the relationship between
@@
 One variable missing from our analysis that we would have liked to
-include is cloud cover.\footnote{We gratefully acknowledge Christopher
-  Adolph for this suggestion.} This variable would have allowed us to
-account for the possibility of brilliant, sunny, but cold winter days,
-as well as warm, dry, but gloomy summer days. Unfortunately, the
-dataset downloaded from Forecast.io included a great deal of missing
-values for this variable, and we were unable to identify another
-suitable source for this data in time for this article.
+include is cloud cover. This variable would have allowed us to account
+for the possibility of brilliant, sunny, but cold winter days, as well
+as warm, dry, but gloomy summer days. Unfortunately, the dataset
+downloaded from Forecast.io included a great deal of missing values
+for this variable, and thus future work might include identification
+of a more suitable source of these data.
 
@@
 Fremont bridge is somewhat representative of a high-volume bicycle
-facility in Seattle, however we might, for example, see a decreased
+facility in Seattle. However we might, for example, see a decreased
 effect size for University of Washington session status if we looked
@@
 next. A future direction would be to consider models, such as ARIMA,
-that more explicitly account for time.
+that more explicitly account for time, though that decision would
+represent a potentially undesirable increase in model complexity.
 
-\section*{Conclusion}
+\chapter*{Conclusions}
 This research set out to help clarify the relationship between weather
@@
 
-For each term included in our model, we found statistically
-significant effects. More importantly, through the use of
-counterfactual simulation, we estimated what we deem to be substantial
-effect sizes associated with each predictor variable.
+For each term included in our model except temperature squared, we
+found statistically significant effects. More importantly, through the
+use of counterfactual simulation, we estimated what we deem to be
+substantial effect sizes associated with each predictor variable.
 
 This article contributes to the existing literature by demonstrating
-the use of a negative binomial model for bicycle counts. In addition,
-the results presented here were generated from data collected over a
-long study period of two years.
+the use of an appropriately chosen negative binomial model for bicycle
+counts. It provides an additional methodological contribution in
+illustrating counterfactual simulation and visualization to create
+more compelling and intuitive results summaries for
+non-statisticians. Finally, the results presented here were generated
+from data collected over a relatively long study period of two years.
 
@@
 policy makers may want to develop strategies that help mitigate the
-impacts of the natural environment during the winter months on the
-increasing number of bicyclists. In other words, the delta between
-warm dry days and cold wet days should be treated as the opportunity
-frontier. Future research could focus on determining what, if any,
-programmatic or built interventions could ameliorate unfavorable cold-
-and wet-weather bicycling conditions.
+impacts of the natural environment during the winter months. In other
+words, the delta between warm dry days and cold wet days should be
+treated as the opportunity frontier. Future research could focus on
+determining what, if any, programmatic or built interventions could
+ameliorate unfavorable cold- and wet-weather bicycling conditions.
+
+\chapter*{Acknowledgment}
+We acknowledge Profs. Elena Erosheva and Christopher Adolph for their
+guidance on the development of this topic. We also acknowledge Giles
+Goetz and Robin Gold for their involvement on a prior incarnation of
+this work.
 

### 2016-02-07 article/article.tex
edits based on reviewer comments

@@
 counter. However, we believe these crossings are rare at this location
-due to the design of the facility, which directs bicyclists to enter
-the sidewalk, and from our own experience riding and observing other
+due to the design of the facility (metal roadway grating and wide
+sidewalks on the approach), which directs bicyclists to enter the
+sidewalk, and from our own experience riding and observing other
 riders. Counts are aggregated into 15 minute intervals by the City of
@@
 programming language along with several add-on
-packages \parencite{Grolemund:2011aa,Wickham:2011aa,Couture-Beil:2014aa,Lang:2014aa,R-Core-Team:2014aa}.\footnote{In
+packages \parencite{Grolemund:2011aa,Wickham:2011aa,Couture-Beil:2014aa,Lang:2014aa,R-Core-Team:2014aa}. In
 addition, the code used to run the analysis, as well as the final
-dataset are available as a github repository: \url{https://github.com/pschmied/bikecounts-mle}}
+dataset are available as a github repository [NEED IN-TEXT CITATION FOR https://github.com/pschmied/bikecounts-mle]
 Bicycle counts were aggregated by day, and then joined to weather data
@@
 Gilman Trail to the University of Washington. We also felt that this
-variable was a suitable proxy for the ``school season,'' which more
+variable was a suitable proxy for the ``school season,'' (i.e. times
+when significant numbers of students are not on holiday) and more
 broadly captures whether or not other local schools are in session.
@@
 
-Inclusion of the holiday variable was an attempt to account for
-some low outlier counts. Upon inspection of the dataset, Christmas and
-Thanksgiving in particular had very low counts of bicycles relative to
-the days preceding and following. Relatedly, but not accounted for by
-any variable in our model, are some of the high outlier counts. Upon
+Inclusion of the holiday variable (as recognized by the University of
+Washington) was an attempt to account for some low outlier
+counts. Upon inspection of the dataset, Christmas and Thanksgiving in
+particular had very low counts of bicycles relative to the days
+preceding and following. Relatedly, but not accounted for by any
+variable in our model, are some of the high outlier counts. Upon
 inspection, some of the highest counts were observed on National Bike
@@
 simplification would be less justifiable in locations that experience
-greater daily variation in precipitation or in locations that have a
+greater daily variation in precipitation, greater variety of
+precipitation types (e.g. snow), or in locations that have a
 predictable pattern of precipitation during certain hours.
@@
 With all other factors held constant we see that, on days when the
-university is in session, we would expect an average of approximately
-367 additional bicycles observed. Similarly, we see a roughly linear
+university is in session, we would expect to observe an average of
+approximately 367 additional bicycles as compared to days when the
+university is not in session. Similarly, we see a roughly linear
 increase in bicycles with increased day length.
@@
 very high temperatures---likely due to Seattle's moderate summer
-climate, and the lack of extremely warm days in our dataset.
+climate, and the lack of extremely warm days in our dataset (85
+degrees Fahrenheit was the maximum observed).
 
@@
 for the possibility of brilliant, sunny, but cold winter days, as well
-as warm, dry, but gloomy summer days. Unfortunately, the dataset
-downloaded from Forecast.io included a great deal of missing values
-for this variable, and thus future work might include identification
-of a more suitable source of these data.
+as warm, dry, but gloomy summer days. On such days, we might observe
+counts otherwise uncharacteristic of the season. Unfortunately, the
+dataset downloaded from Forecast.io included a great deal of missing
+values for this variable, and thus future work might include
+identification of a more suitable source of these data.
 
